### 2007-06-21 article/Article.tex
Beginning of article
darcs-hash:20070621061646-fafb9-b10015bded8627a02ed0e3c7e792de524360f3de.gz

%
% Complete documentation on the extended LaTeX markup used for Insight
% documentation is available in ``Documenting Insight'', which is part
% of the standard documentation for Insight.  It may be found online
% at:
%
%     http://www.itk.org/

\documentclass{InsightArticle}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
%  hyperref should be the last package to be loaded.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[dvips,
bookmarks,
bookmarksopen,
backref,
colorlinks,linkcolor={blue},citecolor={blue},urlcolor={blue},
]{hyperref}
% to be able to use options in graphics
\usepackage{graphicx}
% for pseudo code
\usepackage{listings}
% subfigures
\usepackage{subfigure}

%  This is a template for Papers to the Insight Journal. 
%  It is comparable to a technical report format.

% The title should be descriptive enough for people to be able to find
% the relevant document. 
\title{Efficient implementation of kernel filtering}

% Increment the release number whenever significant changes are made.
% The author and/or editor can define 'significant' however they like.
\release{0.00}

% At minimum, give your name and an email address.  You can include a
% snail-mail address if you like.
\author{Richard Beare{$^1$} {\small{and}} Ga\"etan Lehmann{$^2$}}
\authoraddress{{$^1$}Department of Medicine, Monash University, Australia.\\
{$^2$}INRA, UMR 1198; ENVA; CNRS, FRE 2857, Biologie du D\'eveloppement et
Reproduction, Jouy en Josas, F-78350, France}

\begin{document}
\maketitle

\ifhtml
\chapter*{Front Matter\label{front}}
\fi

\begin{abstract}
\noindent
% The abstract should be a paragraph or two long, and describe the
% scope of the document.
Kernel based filtering is one of the fundamental tools of image
analysis and processing. A number of approaches have been developed
that allow efficient implementation of such filters even when the
kernel size is large. This article reviews some of these methods and
provides introduces their ITK implementations.
\end{abstract}

\tableofcontents

\section{Introduction}
A kernel based filtering process replaces the pixel at the kernel
origin with the result of applying a function to all pixels defined by
the kernel. Many useful filters, including edge detection and gradient
filters, smoothing filters and rank and morphology filters fall into
this category. Direct implementations of such filters typically
involve visiting all pixels defined by the kernel in order to evaluate
the filter function. Such an approach is usually simple to implement
-- in ITK it is made simple by the neighborhood iterators -- but leads
to an algorithm complexity proportional to the number of pixels in the
kernel (or $O(n^d)$, where $n$ is the kernel size and $d$ is the
dimenionality). Such complexity tends to restrict application of such
filters to small kernels.

A number of classical methods exist for reducing this complexity to
more managable, in some cases kernel size independent, levels. ITK
already exploits such techniques for gaussian convolution
operations. This paper describes the classical techniques for
optimized mathematical morphology filters, rank filters and certain
convolution filters. These filters make the use of large kernels,
which can be very useful in many applications, practical on
conventional computing hardware.

\section{Separability and recursive implementations}
The two approaches most typically used to reduce complexity of kernel
based filters are separability and recursive computation, both of
which are used in the ITK implementation of gaussian convolution
filters. A separable filter implements a multidimensional kernel by
cascading several one dimensional kernels, therefore reducing
complexity from $O(n^d)$ to $O(nd)$. The second approach exploits
redundancy that might be present in the computations of kernel
functions at neighboring locations, leading, in some cases, to a complexity
independent of $n$.

\section{Mathematical morphology operations}
Two optimized forms of the mathematical morphology operations of
erosion, dilation, opening and closing are presented here. The kernel
is usually referred to as the structring element in mathematical
morphology. The methods described here are applied to ``flat''
structuring elements, that is structuring elements without
weights. The first method can be applied to arbitary structuring
elements while the second can be applied to line structuring elements.

\subsection{Arbitary structuring elements}
\label{sect:MMmovingHist}
The method described in \cite{Vandroogenbroeck96.3} relies on the
simple concept of an updatable histogram. A histogram is computed for
a kernel located at the first voxel. The histogram at the neighboring
voxel can then be computed by including newly included voxels and
removing newly excluded voxels. The list of included and excluded
voxels corresponding to movement in any direction can be computed when
the structuring element is created, and the direction with the
smallest number of changes should be selected as the direction for
sweeping the kernel across the image. The erosion or dilation at each
location is computed by selecting the minimum or maximum from the
histogram. This approach is very efficient when 8 or 16 bit pixels are
used because the histogram can be represented as an array, with place
holders tracking the current maximum or minimum increasing
performance. More sophisticated histogram representations are
necessary for larger pixel types. Our implementation uses c++ maps.

This methodology reduces the complexity from $O(n^d)$ to $O(n^{d-1})$,
while keeping the structuring element identical to the direct
implementation.

\subsection{Decomposition of structuring elements}
Morphological erosions and dilations are separable -- successive
dilations by orthogonal lines is equivalent to dilation by a rectangle
with sides equal to the line lengths. This means that any
hyper-rectangular structuring element can be constructed using lines
parallel to the image axes.

Approximations of more complex shapes, particulary circles and
ellipses can constructed using a number of lines at evenly spaced
angles \cite{Adams93}. It is difficult to create a structuring element
with a precisely defined radius using this method because line
structuring elements from which the circle structuring element is
composed must have odd length and there are practical limits due to
the realities of underlying digital grid representation of images. In
addition it is possible that the structuring elements may not be truly
translation invariant due to the representation of line used
(e.g. Bresenham) in the decomposition. However, precisely defined
radii are rarely critical when a large structuring element is called
for. An example of a structuring element created using line
structuring elements is shown in Figure \ref{fig:circledecomposition}. Composition of regular
shapes, such as hexagons and octagons is more accurate.

\begin{figure}[htbp]
\centering
\includegraphics{kernel}
\caption{Approximate circular structuring element, radius 25, constructed using 8 lines.\label{fig:circledecomposition}}
\end{figure}

It is also possible, in theory, to construct 3D structuring elements
in similar ways. The construction of a hyper-rectangle is trivial,
however construction of spheres is more problematic. The code
discussed later provides preliminary implementations based on some
platonic solids and various spherical approximateions, but further
testing and development is needed.

\subsection{Line structuring elements}
The decompositions discussed above are important because an efficient,
recursive, implementation of erosion and dilations along lines
exists. This method was introduced in \cite{Gil1993,vanHerk1992a} and
can compute an erosion or dilation in 3 operations per pixel,
independent of structuring element length. \cite{Gil2000} recently
reduced the cost to 1.5 pixels per pixel, but the procedure is more
complicated and there are reports of no speedup in reality.

The original algorithms utilize forward and backward running maxima
(for dilation) for the length of the structuring element from a pixel
of interest. The dilation can then be computed for a region the size
of the structuring element around the point of interest by comparing
values on the running extrema that are separated by the structuring
element length. This is illustrated in Figure \ref{fig:vHGWmethod}

\begin{figure}[htbp]
\centering
\includegraphics[scale=0.5]{vHGWexpl}
\caption{The van Herk, Gil, Werman method.\label{fig:vHGWmethod}}
\end{figure}

There has been a method published recently that offers improved
performance \cite{Vandroogenbroeck2005Morphological} and the ability
to perform a line opening directly (rather than using an
erosion/dilation cascade). This method has been implemented in the
filters discussed later in the article, but performance issues are not
yet clear. The method employs histograms and therefore needs more
complex data structures when applied to higher precision data,
potentially any speed advantage.

\section{Rank filters}
Efficient implemetations of median and rank filters can be carried out
using exactly the same approach as discussed for morphological
operations with arbitary structuring elements. The method was
originally proposed in \cite{Huang79}. The implementation discussed
later supports arbitary kernel shapes and pixel types.

Rank filters are not separable. However the performance benefits
offered by separability make it worth pretending they are. If median
filtering is being used to provide robust noise filtering or
background estimation then a separable approximation is worth
testing. This concept was originally proposed in \cite{Narendra81}.

\appendix

\bibliographystyle{plain}
\bibliography{InsightJournal,local}
\nocite{ITKSoftwareGuide}

\end{document}

### 2007-06-21 itkBoxMeanCalculatorImageFilter.h
compute the number of pixels in the kernel when the iterator is on the border of the image
darcs-hash:20070621100225-2fc9d-50c924096e8369d2b73d4c677fe5611fb0a2cbe2.gz

@@ -46,6 +46,8 @@ public:
                       TOutputImage::ImageDimension);
   itkStaticConstMacro(InputImageDimension, unsigned int,
                       TInputImage::ImageDimension);
+  itkStaticConstMacro(ImageDimension, unsigned int,
+                      TInputImage::ImageDimension);
 
 
 #ifdef ITK_USE_CONCEPT_CHECKING

### 2007-06-21 itkBoxMeanImageFilter.h
drop the threading support, to write a test
darcs-hash:20070621101511-2fc9d-3876a41cf7058a7509d71c94b60e54543a7d7853.gz

@@ -80,9 +80,10 @@ protected:
   void GenerateInputRequestedRegion() ;
   
   /** Multi-thread version GenerateData. */
-  void  ThreadedGenerateData (const OutputImageRegionType& 
-                              outputRegionForThread,
-                              int threadId) ;
+//   void  ThreadedGenerateData (const OutputImageRegionType& 
+//                               outputRegionForThread,
+//                               int threadId);
+  void GenerateData();
 
 
 private:

### 2007-06-22 article/Article.tex
Extended article
darcs-hash:20070622010226-fafb9-8257848d45b237122aee00f0665850732a4de374.gz

@@
 The method described in \cite{Vandroogenbroeck96.3} relies on the
-simple concept of an updatable histogram. A histogram is computed for
-a kernel located at the first voxel. The histogram at the neighboring
-voxel can then be computed by including newly included voxels and
-removing newly excluded voxels. The list of included and excluded
-voxels corresponding to movement in any direction can be computed when
-the structuring element is created, and the direction with the
-smallest number of changes should be selected as the direction for
-sweeping the kernel across the image. The erosion or dilation at each
-location is computed by selecting the minimum or maximum from the
-histogram. This approach is very efficient when 8 or 16 bit pixels are
-used because the histogram can be represented as an array, with place
-holders tracking the current maximum or minimum increasing
-performance. More sophisticated histogram representations are
-necessary for larger pixel types. Our implementation uses c++ maps.
+simple concept of an updatable histogram, and is often described as a
+``moving histogram'' approach. A histogram is computed for a kernel
+located at the first voxel. The histogram at the neighboring voxel can
+then be computed by including newly included voxels and removing newly
+excluded voxels. The list of included and excluded voxels
+corresponding to movement in any direction can be computed when the
+structuring element is created, and the direction with the smallest
+number of changes should be selected as the direction for sweeping the
+kernel across the image. The erosion or dilation at each location is
+computed by selecting the minimum or maximum from the histogram. This
+approach is very efficient when 8 or 16 bit pixels are used because
+the histogram can be represented as an array, with place holders
+tracking the current maximum or minimum increasing performance. More
+sophisticated histogram representations are necessary for larger pixel
+types. Our implementation uses c++ maps.
 
@@
 \subsection{Decomposition of structuring elements}
+\label{sect:MMdecomp}
 Morphological erosions and dilations are separable -- successive
@@
 
-There has been a method published recently that offers improved
+A method published recently that offers improved
 performance \cite{Vandroogenbroeck2005Morphological} and the ability
@@
 originally proposed in \cite{Huang79}. The implementation discussed
-later supports arbitary kernel shapes and pixel types.
+later supports arbitary kernel shapes and pixel types as well as any
+choice of rank.
 
@@
 background estimation then a separable approximation is worth
-testing. This concept was originally proposed in \cite{Narendra81}.
+testing. This concept was originally proposed in \cite{Narendra81}. In
+fact it is also posible to apply the decomposition discussed in
+Section \ref{sect:MMdecomp} to median filters to achieve a closer
+approximation to circular kernels.
+
+\section{Mean and variance filters}
+A simplified version of the moving histogram approach can also be used
+to efficiently compute the mean or variance of an arbitrarily shaped
+kernel. The histogram used for morphology and rank filters is a data
+structure that, once established for one pixel, can be easily updated
+to represent the kernel for a neighboring pixel. The data structure
+required to perform the equivalent function for mean and variance
+operations is much simpler -- all that is necessary is the sum of
+pixel values under the kernel (or sum of squared values for variance
+calculations), and the kernel size. The sum can be updated by adding
+the new values and subtracting the old and the output computed by
+dividing by the kernel size.
+
+A more efficient strategy was proposed in \cite{Crow84} for rectangular
+kernels. This approach first computes an accumulator image in which
+each $i,j$ is repaced by .. This accumulator image may be computed
+recursively using local neighborhood values. The mean of rectangles of
+any size may then be computed using accumulated values at the
+rectangle corners. Figures \ref{fig:accum} amd \ref{fig:accumMean}
+illustrate the 2D case and the update formulas are provided in
+Equations \ref{eq:accum} and \ref{eq:accumMean}.
+
+\begin{eqnarray}
+\label{eq:accum}
+b_{i,j} & = & \sum_{x \leq i, y \leq j} a_{x,y} \\ \nonumber
+	& = & b_{i,j-1} + b_{i-1, j} + a_{i,j} - b_{i-1, j-1}
+\end{eqnarray}
+where $b_{i,j}$ is the accumulator value at location $i,j$ and $a_{i,j}$ is the input image intensity at $i,j$.
+
+\begin{equation}
+\label{eq:accumMean}
+m_{i,j} = \frac{1}{wh} (b_{i+w/2,j+h/2} + b_{i-w/2,j-h/2} - b_{i-w/2,j+h/2} + b_{i+w/2,j-h/2})
+\end{equation}
+where $m_{i,j}$ is the output of the mean filter at location $i,j$.
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.5]{accum2D}
+\caption{Recursive computation of the accumulator image - value at $(i,j)$ computed using Equation \ref{eq:accum}.\label{fig:accum}}
+\end{figure}
+
+\begin{figure}[htbp]
+\centering
+\includegraphics[scale=0.5]{accumMean2d}
+\caption{Computation of mean using the accumulated image - lue at $(i,j)$ computed using Equation \ref{eq:accumMean}.\label{fig:accumMean}}
+\end{figure}
+
+This approach to convolution has been used recently in face detection
+applications \cite{Viola2004}, where rectangular filters of many
+different sizes and shapes were needed.
+
+\section{ITK implementation}
 

### 2007-06-23 itkBoxMeanImageFilter.h
code refactoring and successful multithreading
darcs-hash:20070623120453-fafb9-1a64703503fe197c87dbd140785c177f8bbd62c0.gz

@@ -70,6 +70,7 @@ public:
   itkSetMacro(Radius, RadiusType)
   itkGetConstReferenceMacro(Radius, RadiusType)
 
+  void GenerateInputRequestedRegion() ;
     
 
 protected:
@@ -77,13 +78,14 @@ protected:
   BoxMeanImageFilter();
   ~BoxMeanImageFilter() {};
 
-  void GenerateInputRequestedRegion() ;
-  
+#if 1  
   /** Multi-thread version GenerateData. */
-//   void  ThreadedGenerateData (const OutputImageRegionType& 
-//                               outputRegionForThread,
-//                               int threadId);
+  void  ThreadedGenerateData (const OutputImageRegionType& 
+			      outputRegionForThread,
+			      int threadId);
+#else
   void GenerateData();
+#endif
 
 
 private:

### 2007-06-23 itkBoxUtilities.h

@@ -1,0 +1,340 @@
+#ifndef __itk_BoxUtilities_h
+#define __itk_BoxUtilities_h
+#include "itkProgressReporter.h"
+#include "itkShapedNeighborhoodIterator.h"
+#include "itkImageRegionConstIterator.h"
+#include "itkImageRegionIterator.h"
+#include "itkConnectedComponentAlgorithm.h"
+#include "itkConstantBoundaryCondition.h"
+#include "itkImageRegionIteratorWithIndex.h"
+#include "itkImageRegionIterator.h"
+#include "itkImageRegionConstIterator.h"
+#include "itkOffset.h"
+#include "itkNumericTraits.h"
+#include "itkNeighborhoodAlgorithm.h"
+#include "itkBoxAccumulatorImageFilter.h"
+#include "itkShapedNeighborhoodIterator.h"
+#include "itkZeroFluxNeumannBoundaryCondition.h"
+
+namespace itk {
+
+template <class TInputImage, class TOutputImage>
+void
+BoxAccumulateFunction(typename TInputImage::ConstPointer inputImage, 
+		      typename TOutputImage::Pointer outputImage, 
+		      typename TInputImage::RegionType inputRegion,
+		      typename TOutputImage::RegionType outputRegion,
+		      ProgressReporter &progress)
+{
+  // typedefs
+  typedef TInputImage InputImageType;
+  typedef typename TInputImage::RegionType RegionType ;
+  typedef typename TInputImage::SizeType SizeType ;
+  typedef typename TInputImage::IndexType IndexType ;
+  typedef typename TInputImage::PixelType PixelType ;
+  typedef typename TInputImage::OffsetType OffsetType ;
+  typedef TOutputImage OutputImageType;
+  typedef typename TOutputImage::PixelType OutputPixelType ;
+  typedef typename TInputImage::PixelType InputPixelType ;
+  
+  typedef ImageRegionConstIterator<TInputImage> InputIterator;
+  typedef ImageRegionIterator<TOutputImage> OutputIterator;
+
+  typedef ShapedNeighborhoodIterator<TOutputImage> NOutputIterator;
+  InputIterator inIt( inputImage, inputRegion);
+  typename TInputImage::SizeType kernelRadius;
+  kernelRadius.Fill(1);
+
+  NOutputIterator noutIt( kernelRadius, outputImage, outputRegion);
+  // this iterator is face connected
+  setConnectivityPrevious( &noutIt, false );
+
[…]

### 2007-06-23 article/Article.tex
Fixed border issues, as far as possible.
darcs-hash:20070622074102-fafb9-babf2e36ef40df789d6caa002bb4b6dd4fac545f.gz
started performance section in article
darcs-hash:20070623130246-fafb9-d0d75cfd007716fb0ffc65bd869ce393939d23d4.gz

@@
 sophisticated histogram representations are necessary for larger pixel
-types. Our implementation uses c++ maps.
+types. Our implementation uses c++ maps. This is an extension to the
+original paper.
 
@@
 \section{Rank filters}
+\label{sect:rank}
 Efficient implemetations of median and rank filters can be carried out
@@
 \section{Mean and variance filters}
+\label{sect:meanVar}
 A simplified version of the moving histogram approach can also be used
@@
 \section{ITK implementation}
+The filters discusses in this article are contained in 3 packages --
+{\em consolidatedMorphology}, {\em fastRankMean} and {\em
+boxConvolution}. The most recent versions of these packages can be
+obtained from {\bf link to Gaetan's darcs repo}. All of the packages
+include multithreaded filters and include many examples.
+
+\subsection{Morphology filters -- consolidatedMorphology}
+An earlier version of the {\em consolidatedMorphology} package has
+been submitted to the InsightJournal. However many improvements and
+bug fixes have been included since then.
+
+The morphology filters utilize a new class -- {\em
+FlatStructuringElement} to describe, arbitary and decomposable
+structuring elements. It provides the following methods to create
+structuring elements with different characteristics:
+\begin{itemize}
+  \item \verb$Ball()$ generate a ball structuring element. It takes the 
+	radius of the structuring element as parameter.
+  \item \verb$Box()$ produces a box structuring element. It takes the radius 
+	of the structuring element as parameter.
+  \item \verb$Poly()$ produces an approximation to a circle or sphere using 
+	line structuring elements, taking the desired radius and number of 
+	lines as parameters. In the 3D case an approximation of a sphere will 
+	be attempted. Caution is advised because this is not well tested yet.
+  \item \verb$FromImage()$ produces a structuring element from an image. The image is passed
+as parameter to the method. A optional parameter can also be passed: the pixel value
+to be considered as the foreground value in the image. This value defaults to
+\verb$NumericTraits< PixelType >::max()$. This lets the user visualize the structuring element.
+\end{itemize}
+
+The algorithm used by a filter is set by the {\em SetAlgorithm}
+method, with options of {\em BASIC}, {\em HISTO}, {\em VHGW} and {\em
+ANCHOR} for the traditional ITK implementation, the moving histogram
+algorithm, the van Herk, Gil, Werman algorithm and the anchor
+algorithm respectively. The VHGW and ANCHOR algorithms can only be
+used for Box and Poly structuring elements. VHGW is more thoroughly
+tested.
+
+\subsection{Rank and mean filters -- fastRankMean}
+The {\em fastRankMean} packge using the sliding window method to
+implement rank and mean filters for arbitary kernel shapes. It also
+uses line versions of these kernels to provide a ``separable rank''
+filter -- an fast approximation to a real rank filter as discussed in
+Section \ref{sect:rank}. There are also a number of filters that
+accept a mask input and return the rank or mean value of the
+intersection of kernel and mask. These are experimental and haven't
+been thoroughly tested.
+\section{Performance}
+\subsection{Box Convolution}
+The relative times for optimized and standard convolution are shown in
+Table \ref{tab:perfBoxConv}, and demonstrate significantly improved
+performance sor kernel radii greater than 3. The execution times of
+box convolution should be relatively constant with kernel size, but an
+obvious increase is clear. This is attributable to the process of
+dealing with the boundary conditions, which utilizes shaped
+neighborhood filters. This aspect of the implementation needs further work
+
+\begin{table}[htbp]
+\centering
+\begin{tabular}{ccc}
+\hline
+Size  &  Direct &  Box \\
+\hline
+\hline
+1   &    0.0109048   &    0.0233631 \\
+2   &    0.022008    &    0.0245971 \\
+3   &    0.0434739   &    0.0263161 \\
+4   &    0.0780701   &    0.0283061 \\
+5   &    0.129174    &    0.0310797 \\
+6   &    0.199876    &    0.0344833 \\
+7   &    0.289754    &    0.0382629 \\
+8   &    0.407095    &    0.043121 \\
+9   &    0.552862    &    0.0489335 \\
+10  &    0.728667    &    0.0555441 \\
+15  &    2.17149     & 0.107776 \\
+20  &    4.83458     & 0.199508 \\
+25  &    9.12699     & 0.372637 \\
+\hline
+
+\end{tabular}
+\caption{Execution times for direct and box convolution implementations applied to the $256 \times 256$ cthead image.\label{tab:perfBoxConv}}
+\end{table}
 

### 2007-06-25 article/Article.tex
Mean working on 2D and 3D images.
darcs-hash:20070625033126-fafb9-b47ae2488befbccd4fa7277a1a3ce5329088fcba.gz

@@
 been thoroughly tested.
+
+\subsection{Rectangular convolution -- boxConvolution}
+The {\em boxConvolution} uses the accumulation method described in
+Section \ref{sect:meanVar} to implement rectangular mean and variance
+computations. The filters implement different boundary conditions to
+{\em itkMeanImageFilter}, with the mean of the kernel inside the image
+being computed.
+
 \section{Performance}
+\subsection{Morpholgy}
+
+\subsection{Rank and Mean filters}
+
 \subsection{Box Convolution}
@@
 Table \ref{tab:perfBoxConv}, and demonstrate significantly improved
-performance sor kernel radii greater than 3. The execution times of
-box convolution should be relatively constant with kernel size, but an
-obvious increase is clear. This is attributable to the process of
-dealing with the boundary conditions, which utilizes shaped
-neighborhood filters. This aspect of the implementation needs further work
-
+performance for kernel radii greater than 2. The times indicate the
+theoretical prediction that the complexity of the algorithm is
+independent of kernel size.
 \begin{table}[htbp]
@@
 Size  &  Direct &  Box \\
+1  &     0.014205    &    0.0256313 \\
+2  &     0.0355665   &    0.0257389 \\
+3  &     0.0507895   &    0.0194342 \\
+4  &     0.0745054   &    0.0152253 \\
+5  &     0.124225    &    0.0149731 \\
+6  &     0.187713    &    0.0149678 \\
+7  &     0.27351     &    0.0152494 \\
+8  &     0.411708    &    0.0155574 \\
+9  &     0.518938    &    0.0156469 \\
+10 &     0.682868    &    0.0158462 \\
+15  &    2.02712     &    0.0165146 \\
+20  &    4.46925     &    0.0176763 \\
+25  &    8.43348      & 0.0182798 \\
 \hline
 \hline
-1   &    0.0109048   &    0.0233631 \\
-2   &    0.022008    &    0.0245971 \\
-3   &    0.0434739   &    0.0263161 \\
-4   &    0.0780701   &    0.0283061 \\
-5   &    0.129174    &    0.0310797 \\
-6   &    0.199876    &    0.0344833 \\
-7   &    0.289754    &    0.0382629 \\
-8   &    0.407095    &    0.043121 \\
-9   &    0.552862    &    0.0489335 \\
-10  &    0.728667    &    0.0555441 \\
-15  &    2.17149     & 0.107776 \\
-20  &    4.83458     & 0.199508 \\
-25  &    9.12699     & 0.372637 \\
-\hline
-
 \end{tabular}
-\caption{Execution times for direct and box convolution implementations applied to the $256 \times 256$ cthead image.\label{tab:perfBoxConv}}
+\caption{Execution times for direct and box convolution implementations applied to the $256 \times 256$ cthead image on a ??.\label{tab:perfBoxConv}}
 \end{table}

### 2007-06-25 itkBoxUtilities.h

@@ -4,7 +4,6 @@
 #include "itkShapedNeighborhoodIterator.h"
 #include "itkImageRegionConstIterator.h"
 #include "itkImageRegionIterator.h"
-#include "itkConnectedComponentAlgorithm.h"
 #include "itkConstantBoundaryCondition.h"
 #include "itkImageRegionIteratorWithIndex.h"
 #include "itkImageRegionIterator.h"
@@ -17,6 +16,54 @@
 #include "itkZeroFluxNeumannBoundaryCondition.h"
 
 namespace itk {
+
+template< class TIterator >
+TIterator*
+setConnectivityEarlyBox( TIterator* it, bool fullyConnected=false )
+{
+  // activate the "previous" neighbours
+  typename TIterator::OffsetType offset;
+  it->ClearActiveList();
+  if( !fullyConnected) 
+    {
+    // only activate the neighbors that are face connected
+    // to the current pixel. do not include the center pixel
+    offset.Fill( 0 );
+    for( unsigned int d=0; d < TIterator::Dimension; ++d )
+      {
+      offset[d] = -1;
+      it->ActivateOffset( offset );
+//       offset[d] = 1;
+//       it->ActivateOffset( offset );
+      offset[d] = 0;
+      }
+    }
+  else
+    {
+    // activate all neighbors that are face+edge+vertex
+    // connected to the current pixel. do not include the center pixel
+    unsigned int centerIndex = it->GetCenterNeighborhoodIndex();
+    for( unsigned int d=0; d < centerIndex; d++ )
+      {
+      offset = it->GetOffset( d );
+      // check for positives in any dimension
+      bool keep = true;
+      for (unsigned i = 0; i < TIterator::Dimension; i++)
+	{
+	if (offset[i] > 0)
+	  {
+	  keep = false;
+	  break;
+	  }
+	}
+      if (keep)
+	it->ActivateOffset( offset );
+      }
+    offset.Fill(0);
+    it->DeactivateOffset( offset );
+    }
+  return it;
+}
 
 template <class TInputImage, class TOutputImage>
 void
@@ -46,8 +93,8 @@ BoxAccumulateFunction(typename TInputImage::ConstPointer inputImage,
   kernelRadius.Fill(1);
 
   NOutputIterator noutIt( kernelRadius, outputImage, outputRegion);
-  // this iterator is face connected
[…]

### 2007-06-25 article/Article.tex
Corrections to article
darcs-hash:20070625044023-fafb9-4cc85177df113c5637fc39f7f1a9b7fcf1cdce9b.gz

@@
 analysis and processing. A number of approaches have been developed
-that allow efficient implementation of such filters even when the
-kernel size is large. This article reviews some of these methods and
-provides introduces their ITK implementations.
+over the years that allow efficient implementation of such filters
+even when the kernel size is large. This article reviews some of these
+methods and introduces their ITK implementations.
 \end{abstract}
@@
 involve visiting all pixels defined by the kernel in order to evaluate
-the filter function. Such an approach is usually simple to implement
+the filter function. Such an approach is usually easy to implement
 -- in ITK it is made simple by the neighborhood iterators -- but leads
@@
 kernel (or $O(n^d)$, where $n$ is the kernel size and $d$ is the
-dimenionality). Such complexity tends to restrict application of such
+dimensionality). Such complexity tends to restrict application of such
 filters to small kernels.
@@
 A number of classical methods exist for reducing this complexity to
-more managable, in some cases kernel size independent, levels. ITK
-already exploits such techniques for gaussian convolution
+more manageable, in some cases kernel size independent, levels. ITK
+already exploits such techniques for Gaussian convolution
 operations. This paper describes the classical techniques for
@@
 based filters are separability and recursive computation, both of
-which are used in the ITK implementation of gaussian convolution
+which are used in the ITK implementation of Gaussian convolution
 filters. A separable filter implements a multidimensional kernel by
@@
 erosion, dilation, opening and closing are presented here. The kernel
-is usually referred to as the structring element in mathematical
+is usually referred to as the structuring element in mathematical
 morphology. The methods described here are applied to ``flat''
 structuring elements, that is structuring elements without
-weights. The first method can be applied to arbitary structuring
+weights. The first method can be applied to arbitrary structuring
 elements while the second can be applied to line structuring elements.
 
-\subsection{Arbitary structuring elements}
+\subsection{Arbitrary structuring elements}
 \label{sect:MMmovingHist}
@@
 Morphological erosions and dilations are separable -- successive
-dilations by orthogonal lines is equivalent to dilation by a rectangle
+dilation by orthogonal lines is equivalent to dilation by a rectangle
 with sides equal to the line lengths. This means that any
@@
 
-Approximations of more complex shapes, particulary circles and
+Approximations of more complex shapes, particularly circles and
 ellipses can constructed using a number of lines at evenly spaced
@@
 addition it is possible that the structuring elements may not be truly
-translation invariant due to the representation of line used
-(e.g. Bresenham) in the decomposition. However, precisely defined
+translation invariant due to the representation of line
+(e.g. Bresenham) used in the decomposition. However, precisely defined
 radii are rarely critical when a large structuring element is called
 for. An example of a structuring element created using line
-structuring elements is shown in Figure \ref{fig:circledecomposition}. Composition of regular
-shapes, such as hexagons and octagons is more accurate.
+structuring elements is shown in Figure
+\ref{fig:circledecomposition}. Composition of regular shapes, such as
+hexagons and octagons is more accurate.
 
@@
 discussed later provides preliminary implementations based on some
-platonic solids and various spherical approximateions, but further
+platonic solids and various spherical approximations, but further
 testing and development is needed.
@@
 
-A method published recently that offers improved
-performance \cite{Vandroogenbroeck2005Morphological} and the ability
-to perform a line opening directly (rather than using an
-erosion/dilation cascade). This method has been implemented in the
-filters discussed later in the article, but performance issues are not
-yet clear. The method employs histograms and therefore needs more
-complex data structures when applied to higher precision data,
-potentially any speed advantage.
+A method published recently that offers improved performance
+\cite{Vandroogenbroeck2005Morphological} and the ability to perform a
+line opening directly (rather than using an erosion/dilation
+cascade). This method, known as {\em anchor morphology} has been
+implemented in the filters discussed later in the article, but
+performance issues are not yet clear. The method employs histograms
+and therefore needs more complex data structures when applied to
+higher precision data, potentially reducing any speed advantage.
 
@@
 \label{sect:rank}
-Efficient implemetations of median and rank filters can be carried out
+Efficient implementations of median and rank filters can be carried out
 using exactly the same approach as discussed for morphological
-operations with arbitary structuring elements. The method was
+operations with arbitrary structuring elements. The method was
 originally proposed in \cite{Huang79}. The implementation discussed
-later supports arbitary kernel shapes and pixel types as well as any
+later supports arbitrary kernel shapes and pixel types as well as any
 choice of rank.
@@
 testing. This concept was originally proposed in \cite{Narendra81}. In
-fact it is also posible to apply the decomposition discussed in
+fact it is also possible to apply the decomposition discussed in
 Section \ref{sect:MMdecomp} to median filters to achieve a closer
@@
 
-A more efficient strategy was proposed in \cite{Crow84} for rectangular
-kernels. This approach first computes an accumulator image in which
-each $i,j$ is repaced by .. This accumulator image may be computed
-recursively using local neighborhood values. The mean of rectangles of
-any size may then be computed using accumulated values at the
-rectangle corners. Figures \ref{fig:accum} amd \ref{fig:accumMean}
-illustrate the 2D case and the update formulas are provided in
-Equations \ref{eq:accum} and \ref{eq:accumMean}.
+A more efficient strategy was proposed in \cite{Crow84} for
+rectangular kernels. This approach first computes an accumulator image
+in which each $i,j$ is replaced by the sum of voxels ``left'' and
+``above'' it. This accumulator image may be computed recursively using
+local neighborhood values. The mean of rectangles of any size may then
+be computed using accumulated values at the rectangle corners. Figures
+\ref{fig:accum} and \ref{fig:accumMean} illustrate the 2D case and the
+update formulas are provided in Equations \ref{eq:accum} and
+\ref{eq:accumMean}.
 
@@
 \includegraphics[scale=0.5]{accumMean2d}
-\caption{Computation of mean using the accumulated image - lue at $(i,j)$ computed using Equation \ref{eq:accumMean}.\label{fig:accumMean}}
+\caption{Computation of mean using the accumulated image - value at $(i,j)$ computed using Equation \ref{eq:accumMean}.\label{fig:accumMean}}
 \end{figure}
@@
 
+The availability of line morphology operations means that new
+operations, such as ``union of openings'' can be implemented easily.
+
 \subsection{Rank and mean filters -- fastRankMean}
-The {\em fastRankMean} packge using the sliding window method to
+The {\em fastRankMean} packge uses the sliding window method to
 implement rank and mean filters for arbitary kernel shapes. It also
@@
 {\em itkMeanImageFilter}, with the mean of the kernel inside the image
-being computed.
+being computed. The filters in this package are multithreaded and deal
+with arbitary dimensions.
 
@@
 
+\section{Comments and Conclusions}
+
 \appendix

### 2007-06-25 itkBoxAccumulatorImageFilter.h
drop unused Set/GetRadius() methods
darcs-hash:20070625121623-2fc9d-b3b6469c49920b474f99cb9f677315b46e5aa564.gz

@@ -54,10 +54,6 @@ public:
   /** n-dimensional Kernel radius. */
   typedef typename TInputImage::SizeType RadiusType ;
 
-  // itkSetMacro(Radius, RadiusType);
-  void SetRadius( const RadiusType );
-  itkGetMacro(Radius, RadiusType);
-
   void GenerateData();
 
 protected:

### 2007-06-25 itkBoxUtilities.h
iterators shouldn't be incremented twice in a loop
darcs-hash:20070625130326-2fc9d-816804315967baaf36fcf836989ede22bd988c58.gz

@@ -57,7 +57,10 @@ setConnectivityEarlyBox( TIterator* it, bool fullyConnected=false )
 	  }
 	}
       if (keep)
+        {
+//         std::cout << "+  " << offset << std::endl;
 	it->ActivateOffset( offset );
+        }
       }
     offset.Fill(0);
     it->DeactivateOffset( offset );
@@ -121,6 +124,7 @@ BoxAccumulateFunction(typename TInputImage::ConstPointer inputImage,
       if (offset[k] != 0)
 	w *= offset[k];
       }
+//     std::cout << offset << "  " << w << std::endl;
     Weights.push_back(w);
     }
 
@@ -133,8 +137,6 @@ BoxAccumulateFunction(typename TInputImage::ConstPointer inputImage,
       Sum += sIt.Get() * Weights[k];
       }
     noutIt.SetCenterPixel(Sum + inIt.Get());
-    ++inIt;
-    ++noutIt;
     progress.CompletedPixel();
     }
 }

### 2007-06-25 itkBoxMeanCalculatorImageFilter.h
fix threading support
darcs-hash:20070625133739-2fc9d-529b76c36442f2e49a57ea820f14105238535f29.gz

@@ -73,7 +73,9 @@ protected:
 
 //  void GenerateInputRequestedRegion() ;
   
-  void  GenerateData ();
+  void  ThreadedGenerateData (const OutputImageRegionType& 
+			      outputRegionForThread,
+			      int threadId);
 
 
 private:

### 2007-06-25 itkBoxMeanImageFilter.h

@@ -78,7 +78,7 @@ protected:
   BoxMeanImageFilter();
   ~BoxMeanImageFilter() {};
 
-#if 1  
+#if 0  
   /** Multi-thread version GenerateData. */
   void  ThreadedGenerateData (const OutputImageRegionType& 
 			      outputRegionForThread,

### 2007-06-27 itkBoxMeanImageFilter.h
restore Richard's threading code, and fix the bug in region limit computation which was making the filter crash
darcs-hash:20070627103912-2fc9d-6934eebed8235da411a503a6cdcb6d4c67b69a59.gz

@@ -78,7 +78,7 @@ protected:
   BoxMeanImageFilter();
   ~BoxMeanImageFilter() {};
 
-#if 0  
+#if 1
   /** Multi-thread version GenerateData. */
   void  ThreadedGenerateData (const OutputImageRegionType& 
 			      outputRegionForThread,

### 2007-06-27 itkBoxUtilities.h

@@ -208,7 +208,7 @@ BoxMeanCalculatorFunction(typename TInputImage::ConstPointer accImage,
     {
     kernelSize[i] = Radius[i] * 2 + 1;
     internalRadius[i] = Radius[i] + 1;
-    RegionLimit[i] = inputRegion.GetSize()[i] - RegionStart[i] - 1;
+    RegionLimit[i] = inputRegion.GetSize()[i] + RegionStart[i] - 1;
     }
 
   typedef typename itk::NumericTraits<OutputPixelType>::RealType AccPixType;

### 2007-06-27 itkBoxUtilities.h
make BoxAccumulateFunction take a functor argument
darcs-hash:20070627144902-2fc9d-5566cea487aa8d00bec8c879b4f7bf0599affaea.gz

@@ -14,6 +14,7 @@
 #include "itkBoxAccumulatorImageFilter.h"
 #include "itkShapedNeighborhoodIterator.h"
 #include "itkZeroFluxNeumannBoundaryCondition.h"
+#include "vcl_functional.h"
 
 namespace itk {
 
@@ -68,7 +69,33 @@ setConnectivityEarlyBox( TIterator* it, bool fullyConnected=false )
   return it;
 }
 
-template <class TInputImage, class TOutputImage>
+
+
+namespace Functor {
+
+template<class TValue>
+class ITK_EXPORT Identity
+{
+public:
+  inline const TValue & operator()( const TValue & v )
+    {
+    return v;
+    }
+};
+
+template<class TValue>
+class ITK_EXPORT Square
+{
+public:
+  inline TValue operator()( const TValue & v )
+    {
+    return v*v;
+    }
+};
+
+}
+
+template <class TInputImage, class TOutputImage, class TFunctor >
 void
 BoxAccumulateFunction(typename TInputImage::ConstPointer inputImage, 
 		      typename TOutputImage::Pointer outputImage, 
@@ -128,13 +155,15 @@ BoxAccumulateFunction(typename TInputImage::ConstPointer inputImage,
     Weights.push_back(w);
     }
 
+  TFunctor function;
+
   for (inIt.GoToBegin(), noutIt.GoToBegin(); !noutIt.IsAtEnd(); ++inIt, ++noutIt )
     {
     OutputPixelType Sum = 0;
     int k;
     for (k = 0, sIt = noutIt.Begin(); !sIt.IsAtEnd();++sIt, ++k)
       {
-      Sum += sIt.Get() * Weights[k];
+      Sum += function( sIt.Get() ) * Weights[k];
       }
     noutIt.SetCenterPixel(Sum + inIt.Get());
     progress.CompletedPixel();

### 2007-06-27 itkBoxUtilities.h
UNDO: make BoxAccumulateFunction take a functor argument
darcs-hash:20070627144902-2fc9d-0329a788ca2eb6059ccec7e8af0eaa378bea7db5.gz

@@ -14,7 +14,6 @@
 #include "itkBoxAccumulatorImageFilter.h"
 #include "itkShapedNeighborhoodIterator.h"
 #include "itkZeroFluxNeumannBoundaryCondition.h"
-#include "vcl_functional.h"
 
 namespace itk {
 
@@ -69,33 +68,7 @@ setConnectivityEarlyBox( TIterator* it, bool fullyConnected=false )
   return it;
 }
 
-
-
-namespace Functor {
-
-template<class TValue>
-class ITK_EXPORT Identity
-{
-public:
-  inline const TValue & operator()( const TValue & v )
-    {
-    return v;
-    }
-};
-
-template<class TValue>
-class ITK_EXPORT Square
-{
-public:
-  inline TValue operator()( const TValue & v )
-    {
-    return v*v;
-    }
-};
-
-}
-
-template <class TInputImage, class TOutputImage, class TFunctor >
+template <class TInputImage, class TOutputImage>
 void
 BoxAccumulateFunction(typename TInputImage::ConstPointer inputImage, 
 		      typename TOutputImage::Pointer outputImage, 
@@ -155,15 +128,13 @@ BoxAccumulateFunction(typename TInputImage::ConstPointer inputImage,
     Weights.push_back(w);
     }
 
-  TFunctor function;
-
   for (inIt.GoToBegin(), noutIt.GoToBegin(); !noutIt.IsAtEnd(); ++inIt, ++noutIt )
     {
     OutputPixelType Sum = 0;
     int k;
     for (k = 0, sIt = noutIt.Begin(); !sIt.IsAtEnd();++sIt, ++k)
       {
-      Sum += function( sIt.Get() ) * Weights[k];
+      Sum += sIt.Get() * Weights[k];
       }
     noutIt.SetCenterPixel(Sum + inIt.Get());
     progress.CompletedPixel();

### 2007-06-27 itkBoxSigmaImageFilter.h
begin implementation of sigma calculator
darcs-hash:20070627151835-2fc9d-80ece5273c0710adaca16c1bf9b7c92477961f0c.gz

@@ -1,0 +1,105 @@
+#ifndef __itkBoxSigmaImageFilter_h
+#define __itkBoxSigmaImageFilter_h
+
+#include "itkImageToImageFilter.h"
+
+namespace itk {
+
+/**
+ * \class BoxSigmaImageFilter
+ * \brief Implements a fast rectangular mean filter using the
+ * accumulator approach
+ * \author Richard Beare
+ */
+
+template<class TInputImage, class TOutputImage=TInputImage>
+class ITK_EXPORT BoxSigmaImageFilter : 
+    public ImageToImageFilter<TInputImage, TOutputImage>
+{
+public:
+  /** Standard class typedefs. */
+  typedef BoxSigmaImageFilter Self;
+  typedef ImageToImageFilter<TInputImage, TOutputImage>  Superclass;
+  typedef SmartPointer<Self>        Pointer;
+  typedef SmartPointer<const Self>  ConstPointer;
+  
+  /** Standard New method. */
+  itkNewMacro(Self);  
+
+  /** Runtime information support. */
+  itkTypeMacro(BoxSigmaImageFilter, 
+               BoxSigmaImageFilter);
+  
+  /** Image related typedefs. */
+  typedef TInputImage InputImageType;
+  typedef TOutputImage OutputImageType;
+  typedef typename TInputImage::RegionType RegionType ;
+  typedef typename TInputImage::SizeType SizeType ;
+  typedef typename TInputImage::IndexType IndexType ;
+  typedef typename TInputImage::PixelType PixelType ;
+  typedef typename TInputImage::OffsetType OffsetType ;
+  typedef typename Superclass::OutputImageRegionType OutputImageRegionType;
+  typedef typename TOutputImage::PixelType OutputPixelType ;
+
+  /** Image related typedefs. */
+  itkStaticConstMacro(OutputImageDimension, unsigned int,
+                      TOutputImage::ImageDimension);
+  itkStaticConstMacro(InputImageDimension, unsigned int,
+                      TInputImage::ImageDimension);
+
+
+#ifdef ITK_USE_CONCEPT_CHECKING
+  /** Begin concept checking */
+  itkConceptMacro(SameDimension,
+                  (Concept::SameDimension<itkGetStaticConstMacro(InputImageDimension),itkGetStaticConstMacro(OutputImageDimension)>));
+
+  
+  /** End concept checking */
+#endif
+
[…]

### 2007-06-27 itkBoxUtilities.h

@@ -381,6 +381,189 @@ BoxMeanCalculatorFunction(typename TInputImage::ConstPointer accImage,
     }
 }
 
+
+template <class TInputImage, class TOutputImage>
+void
+BoxSigmaCalculatorFunction(typename TInputImage::ConstPointer accImage, 
+			  typename TInputImage::ConstPointer accImage, 
+			  typename TOutputImage::Pointer outputImage, 
+			  typename TInputImage::RegionType inputRegion,
+			  typename TOutputImage::RegionType outputRegion,
+			  typename TInputImage::SizeType Radius,
+			  ProgressReporter &progress)
+{
+  // typedefs
+  typedef TInputImage InputImageType;
+  typedef typename TInputImage::RegionType RegionType ;
+  typedef typename TInputImage::SizeType SizeType ;
+  typedef typename TInputImage::IndexType IndexType ;
+  typedef typename TInputImage::PixelType PixelType ;
+  typedef typename TInputImage::OffsetType OffsetType ;
+  typedef TOutputImage OutputImageType;
+  typedef typename TOutputImage::PixelType OutputPixelType ;
+  typedef typename TInputImage::PixelType InputPixelType ;
+   // use the face generator for speed
+  typedef typename itk::NeighborhoodAlgorithm::ImageBoundaryFacesCalculator<InputImageType> FaceCalculatorType;
+  typedef typename FaceCalculatorType::FaceListType FaceListType;
+  typedef typename FaceCalculatorType::FaceListType::iterator FaceListTypeIt;
+  FaceCalculatorType faceCalculator;
+
+  FaceListType faceList;
+  FaceListTypeIt fit;
+  ZeroFluxNeumannBoundaryCondition<TInputImage> nbc;
+
+  // this process is actually slightly asymmetric because we need to
+  // subtract rectangles that are next to our kernel, not overlapping it
+  SizeType kernelSize, internalRadius, BoxSize, RegionLimit;
+  IndexType RegionStart = inputRegion.GetIndex();
+  for( int i=0; i<TInputImage::ImageDimension; i++ )
+    {
+    kernelSize[i] = Radius[i] * 2 + 1;
+    internalRadius[i] = Radius[i] + 1;
+    RegionLimit[i] = inputRegion.GetSize()[i] + RegionStart[i] - 1;
+    }
+
+  typedef typename itk::NumericTraits<OutputPixelType>::RealType AccPixType;
[…]

### 2007-06-28 article/Article.tex
Acknowledgments
darcs-hash:20070625145545-2fc9d-2f35fef9e2ea60e348efd5fdc10d0e9d23673388.gz
make BoxAccumulateFunction take a functor argument
darcs-hash:20070627144902-2fc9d-5566cea487aa8d00bec8c879b4f7bf0599affaea.gz
add sigma filter - a test is still missing
darcs-hash:20070628103023-2fc9d-fd881259703ac3509863058b153c8a0fa7837cef.gz

@@
 
-\section{Mean and variance filters}
+\section{Mean and standard deviation filters}
 \label{sect:meanVar}
 A simplified version of the moving histogram approach can also be used
-to efficiently compute the mean or variance of an arbitrarily shaped
+to efficiently compute the mean or standard deviation of an arbitrarily shaped
 kernel. The histogram used for morphology and rank filters is a data
@@
 \ref{fig:accum} and \ref{fig:accumMean} illustrate the 2D case and the
-update formulas are provided in Equations \ref{eq:accum} and
+update formulas are provided in Equations \ref{eq:accum}, \ref{eq:accSum}, and
 \ref{eq:accumMean}.
@@
 \begin{equation}
+\label{eq:accSum}
+S_{i,j} = b_{i+w/2,j+h/2} + b_{i-w/2,j-h/2} - b_{i-w/2,j+h/2} + b_{i+w/2,j-h/2}
+\end{equation}
+where $S_{i,j}$ is the sum of the pixel values in the neighborhood centered at $i,j$.
+
+\begin{equation}
 \label{eq:accumMean}
-m_{i,j} = \frac{1}{wh} (b_{i+w/2,j+h/2} + b_{i-w/2,j-h/2} - b_{i-w/2,j+h/2} + b_{i+w/2,j-h/2})
+m_{i,j} = \frac{1}{w.h}S_{i,j}
 \end{equation}
 where $m_{i,j}$ is the output of the mean filter at location $i,j$.
+
+Using the same strategy, the standard deviation can be efficiently computed.
+It requires an accumulator image of the squared pixels values, which can be
+efficiently computed at the same time than the accumulator image. The update
+formulas are provided in Equations \ref{eq:accum}, \ref{eq:accSum},
+\ref{eq:accum2}, \ref{eq:accSum2} and \ref{eq:accumSigma}.
+
+\begin{eqnarray}
+\label{eq:accum2}
+b2_{i,j} & = & \sum_{x \leq i, y \leq j} a_{x,y}^2 \\ \nonumber
+	& = & b2_{i,j-1} + b2_{i-1, j} + a_{i,j}^2 - b2_{i-1, j-1}
+\end{eqnarray}
+where $b2_{i,j}$ is the accumulator value at location $i,j$ for the square image, and $a_{i,j}$ is the input image intensity at $i,j$.
+
+\begin{equation}
+\label{eq:accSum2}
+S2_{i,j} = b2_{i+w/2,j+h/2} + b2_{i-w/2,j-h/2} - b2_{i-w/2,j+h/2} + b2_{i+w/2,j-h/2}
+\end{equation}
+where $S2_{i,j}$ is the sum of the squared pixel values in the neighborhood centered at $i,j$.
+
+\begin{equation}
+\label{eq:accumSigma}
+\sigma_{i,j} = \sqrt{\frac{S2_{i,j} - \frac{S_{i,j}^2}{w.h}}{w.h-1}}
+\end{equation}
+where $\sigma_{i,j}$ is the output of the standard deviation filter at location $i,j$.
 
@@
 
+\section{Acknowledgments}
+We thank Dr Pierre Adenot and MIMA2 confocal facilities
+(\url{http://mima2.jouy.inra.fr}) for providing the 3D test image.
+
 \appendix

### 2007-06-28 itkBoxUtilities.h

@@ -100,7 +100,7 @@ BoxAccumulateFunction(typename TInputImage::ConstPointer inputImage,
   setConnectivityEarlyBox( &noutIt, true );
 
   ConstantBoundaryCondition<OutputImageType> oBC;
-  oBC.SetConstant(0);
+  oBC.SetConstant( NumericTraits< OutputPixelType >::Zero );
   noutIt.OverrideBoundaryCondition(&oBC);
   // This uses several iterators. An alternative and probably better
   // approach would be to copy the input to the output and convolve
@@ -356,7 +356,6 @@ BoxMeanCalculatorFunction(typename TInputImage::ConstPointer accImage,
 template <class TInputImage, class TOutputImage>
 void
 BoxSigmaCalculatorFunction(typename TInputImage::ConstPointer accImage, 
-			  typename TInputImage::ConstPointer accImage, 
 			  typename TOutputImage::Pointer outputImage, 
 			  typename TInputImage::RegionType inputRegion,
 			  typename TOutputImage::RegionType outputRegion,
@@ -373,6 +372,7 @@ BoxSigmaCalculatorFunction(typename TInputImage::ConstPointer accImage,
   typedef TOutputImage OutputImageType;
   typedef typename TOutputImage::PixelType OutputPixelType ;
   typedef typename TInputImage::PixelType InputPixelType ;
+  typedef typename InputPixelType::ValueType ValueType ;
    // use the face generator for speed
   typedef typename itk::NeighborhoodAlgorithm::ImageBoundaryFacesCalculator<InputImageType> FaceCalculatorType;
   typedef typename FaceCalculatorType::FaceListType FaceListType;
@@ -457,14 +457,18 @@ BoxSigmaCalculatorFunction(typename TInputImage::ConstPointer accImage,
       for (oIt.GoToBegin(); !oIt.IsAtEnd(); ++oIt)
         {
 	AccPixType Sum = 0;
+	AccPixType SquareSum = 0;
 	// check each corner
 	for (unsigned k = 0; k < CornerItVec.size(); k++)
 	  {
-	  Sum += Weights[k] * CornerItVec[k].Get();
+          const InputPixelType & i = CornerItVec[k].Get();
+	  Sum += Weights[k] * i[0];
+	  SquareSum += Weights[k] * i[1];
 	  // increment each corner iterator
 	  ++(CornerItVec[k]);
 	  }
-	oIt.Set(static_cast<OutputPixelType>(Sum/pixelscount));
+
[…]

### 2007-06-29 article/Article.tex
Updated article and included some timing
darcs-hash:20070629053316-fafb9-435d466604b8e42b77976421ff1b9202efbf9049.gz

@@
 sophisticated histogram representations are necessary for larger pixel
-types. Our implementation uses c++ maps. This is an extension to the
+types. Our implementation uses c++ maps, which is an extension to the
 original paper.
@@
 with sides equal to the line lengths. This means that any
-hyper-rectangular structuring element can be constructed using lines
-parallel to the image axes.
+hyper-rectangular structuring element can be constructed using several
+orthogonal lines, typically parallel to the axes.
 
@@
 reduced the cost to 1.5 pixels per pixel, but the procedure is more
-complicated and there are reports of no speedup in reality.
+complicated and there are reports of no speedup in practice.
 
@@
 
-A method published recently that offers improved performance
-\cite{Vandroogenbroeck2005Morphological} and the ability to perform a
-line opening directly (rather than using an erosion/dilation
-cascade). This method, known as {\em anchor morphology} has been
+A method known as {\em anchor morphology} has been published recently
+\cite{Vandroogenbroeck2005Morphological} that offers improved
+performance and the ability to perform a line opening directly (rather
+than using an erosion/dilation cascade). This method has been
 implemented in the filters discussed later in the article, but
@@
 
-Using the same strategy, the standard deviation can be efficiently computed.
-It requires an accumulator image of the squared pixels values, which can be
-efficiently computed at the same time than the accumulator image. The update
-formulas are provided in Equations \ref{eq:accum}, \ref{eq:accSum},
+Using the same strategy, the standard deviation can be efficiently
+computed.  It requires an accumulator image of the squared pixels
+values, which can be efficiently computed at the same time as the
+original accumulator image. The update formulas are provided in
+Equations \ref{eq:accum}, \ref{eq:accSum},
 \ref{eq:accum2}, \ref{eq:accSum2} and \ref{eq:accumSigma}.
@@
 \includegraphics[scale=0.5]{accum2D}
-\caption{Recursive computation of the accumulator image - value at $(i,j)$ computed using Equation \ref{eq:accum}.\label{fig:accum}}
+\caption{Recursive computation of the accumulator image - value at $(i,j)$ computed using Equation \ref{eq:accum}.
+The double hashed region is subtracted twice because it is included in both $b_{i,j-1}$ and $b_{i-1, j}$.\label{fig:accum}}
 \end{figure}
@@
 applications \cite{Viola2004}, where rectangular filters of many
-different sizes and shapes were needed.
+different sizes and shapes were needed and could be computed from the same accumulation image.
 
@@
 used for Box and Poly structuring elements. VHGW is more thoroughly
-tested.
+tested. This HISTO algorithm produces the identical results to BASIC,
+and is always faster. It is the default and should be used if you need
+precisely defined, non rectangular, structuring elements. If a Box
+structuring element is needed then VHGW or ANCHOR should be used, as
+they offer the best performance and are exact in that case. Poly
+structuring elements are approximations of circles that will offer
+improved performance with VHGW or ANCHOR algorithms, but the user will
+need to consider whether the approximation is suitable for the
+application in question.
 
@@
 intersection of kernel and mask. These are experimental and haven't
-been thoroughly tested.
+been thoroughly tested. The filters are {\em itkRankImageFilter,
+itkMovingWindowMeanImageFilter, itkFastApproxRankImageFilter,
+itkFastApproxMaskRankImageFilter} and {\em
+itkMaskedRankImageFilter}. itkRankImageFilter implements arbitary
+shaped kernels, itkFastApproxRankImageFilter implements the separable
+approximation and itkMovingWindowMeanImageFilter implements an
+arbitary shaped kernel mean.
 
@@
 being computed. The filters in this package are multithreaded and deal
-with arbitary dimensions.
+with arbitary dimensions. The current filters do not support the mode
+of operation used in face recognition application in which an
+accumulation image is repeatedly ``queried'' for mean values of
+various sized kernels. However introducing such a capability is
+relatively easy.
 
@@
 \subsection{Rank and Mean filters}
+Relative times for sliding window median and mean filters and
+separable median filters are shown in Table \ref{tab:perfRank}. The
+predicted linear complexity is observed for the sliding window
+approaches (complexity of direct approach is $O(n^2)$ in 2D, reduced
+to $O(n)$ by using the sliding window. The separable median exhibits a
+runtime independent of kernel size, as expected.
+\begin{table}[htbp]
+\centering
+\begin{tabular}{cccccc}
+\hline
+Size  &  Direct Median &  Sliding Median & Separable Median & Direct Mean & Sliding Mean\\
+1     &   0.0352 & 0.0341 & 0.0378 & 0.0171 & 0.0229\\
+5     &   0.409  & 0.0823  & 0.0201 & 0.24  &  0.0601 \\
+10    &   1.03   & 0.131 & 0.0193 & 0.803 &  0.0569 \\
+15    &   2.82   & 0.106  & 0.0192 & 2.13  &  0.0819 \\
+20    &   5.85   & 0.138  & 0.0358 & 4.76  &  0.106 \\
+40    &   39.7   & 0.2563  & 0.0198 & 36.6  &  0.203 \\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for median and mean filters using direct and sliding window implementations applied to the $256 \times 256$ cthead image on a 1.8GHz AMD Athlon, 512KB cache.
+The separable median is an approximation and therefore doesn't produce the same result as the direct or sliding algorithms.\label{tab:perfRank}}
+\end{table}
 
@@
 Table \ref{tab:perfBoxConv}, and demonstrate significantly improved
-performance for kernel radii greater than 2. The times indicate the
+performance for kernel radii greater than 1. The times support the
 theoretical prediction that the complexity of the algorithm is
@@
 \centering
-\begin{tabular}{ccc}
+\begin{tabular}{ccccc}
 \hline
-Size  &  Direct &  Box \\
-1  &     0.014205    &    0.0256313 \\
-2  &     0.0355665   &    0.0257389 \\
-3  &     0.0507895   &    0.0194342 \\
-4  &     0.0745054   &    0.0152253 \\
-5  &     0.124225    &    0.0149731 \\
-6  &     0.187713    &    0.0149678 \\
-7  &     0.27351     &    0.0152494 \\
-8  &     0.411708    &    0.0155574 \\
-9  &     0.518938    &    0.0156469 \\
-10 &     0.682868    &    0.0158462 \\
-15  &    2.02712     &    0.0165146 \\
-20  &    4.46925     &    0.0176763 \\
-25  &    8.43348      & 0.0182798 \\
+Size  &  Direct Mean &  Box Mean & Direct sigma & Box sigma\\
+1    &   0.0142 & 0.0226 & 0.0176 & 0.0332 \\
+2    &   0.0279 & 0.0176 & 0.0294 & 0.0258 \\
+3    &   0.0425 & 0.013  & 0.0447 & 0.0193 \\
+4    &   0.077  & 0.0132 & 0.0803 & 0.0195 \\
+5    &   0.128  & 0.0132 & 0.129  & 0.0195 \\
+6    &   0.195  & 0.0134 & 0.195  & 0.0198 \\
+7    &   0.275  & 0.0136 & 0.283  & 0.0198 \\
+8    &   0.392  & 0.0148 & 0.405  & 0.0204 \\
+9    &   0.54   & 0.014  & 0.548  & 0.0201 \\
+10   &   0.706  & 0.0139 & 0.713  & 0.0202 \\
+15   &   2.12   & 0.0158 & 2.18   & 0.021 \\
+20   &   4.62   & 0.0157 & 4.76   & 0.0217 \\
+25   &   8.63   & 0.0164 & 8.91   & 0.0222 \\
 \hline
@@
 \end{tabular}
-\caption{Execution times for direct and box convolution implementations applied to the $256 \times 256$ cthead image on a ??.\label{tab:perfBoxConv}}
+\caption{Execution times in seconds for mean and standard deviation using direct and box convolution implementations applied to the $256 \times 256$ cthead image on a 1.8GHz AMD Athlon, 512KB cache.\label{tab:perfBoxConv}}
 \end{table}
+\subsection{Timing notes}
+You may have noticed that the optimized filters of radius 1 seem to
+regularly perform slower than their larger counterparts. We can think
+of no reason for this. The algorithms that have kernel size
+independent complexity should not be slower for small kernels than
+large ones because the only difference occurs at borders and when the
+sliding window is being created, and both situations favour small
+kernels. We can only guess that there are caching issues involved.
+
+\section{Implementation Notes}
+A lot of effort has been spent implementing and optimizing these filters 
+while attempting to maintain good ITK style. This section summarizes some observations
+and experiences:
+\subsection{Moving histograms}
+The most obvious way of implementing moving histograms involves
+neighborhood iterators.  However the resulting performance wasn't good
+-- it appears that neighborhood iterator complexity is proportional to
+radius rather than the number of active neighbors. A method using
+lists of offsets combined with image Get/SetPixel methods approach was
+used instead.
+
+The original paper recommended moving the window in a zig-zag pattern, i.e horizontally 
+across a the first row, down a step to the second row and then back, but this involves
+accessing data in a reverse raster order which can potentially reduce cache performance.
+An alternative has been implemented that always moves the window in forward raster direction,
+at the cost of additional histogram copies, has been implemented and exhibits improved
+performance for large images. Differences for small images weren't detectable.
+
+\subsection{16 bit data types}
+Histograms for 16 bit data types can be implemented using arrays or maps. One would
+expect arrays to offer a simpler and therefore faster implementation, but our results
+have been mixed. Presumably the performance depends significantly on voxel statistics.
+
+\subsection{Line Operations}
+The van Herk/Gil and Werman and anchor methods operate on image transects, potentially
+at any angle. The transect data needs to be extracted from the image and the processed
+transect written back, with successive parallel transects being processed. 
+Existing iterators don't do this particularly efficiently, so image Get/SetPixel operations 
+are being used. This approach isn't ideal either and needs to be worked on. In general
+there is no need for boundary checks because the intersection between transect and image region
+is calculated, so a potentially random access method without boundary checks would be ideal.
+
+Building dimension independent code also adds complexity to this filter which can be avoided in
+dimension specific implementations. The process of sweeping the transect across the image region
+is much more complex (and therefore slower) when written in a dimension independent fashion.
+
+The standard ITK threading approach isn't ideal for this style of filtering either -- a better 
+option would be splitting the image based on transects instead of blocks. Obviously this
+sort of change isn't practical.
+
+\subsection{Refactoring}
+The three packages do share some components, and some refactoring is
+called for. This most widely used components are histograms and
+sliding window infrastructure, which need to be consolidated.
 
 \section{Comments and Conclusions}
-
-\section{Acknowledgments}
-We thank Dr Pierre Adenot and MIMA2 confocal facilities
-(\url{http://mima2.jouy.inra.fr}) for providing the 3D test image.
+This article has provided background theory for and implementations of
+a number of important approaches to kernel based filtering. These
+methods significantly reduce complexity and execution time, by some
+orders of magnitude in many cases. These approaches are all well
+established in the literature but weren't available in
+ITK. Availability of filtering algorithms which are always fast,
+irrespective of the kernel size, can make much simpler approaches to
+many problems practical.
 

### 2007-06-29 article/Article.tex
spell check
darcs-hash:20070629054141-fafb9-f6a798c062900ad4c9f5790a2f7d65cd1425f14f.gz

@@
 The method described in \cite{Vandroogenbroeck96.3} relies on the
-simple concept of an updatable histogram, and is often described as a
+simple concept of an up-datable histogram, and is often described as a
 ``moving histogram'' approach. A histogram is computed for a kernel
@@
 obtained from {\bf link to Gaetan's darcs repo}. All of the packages
-include multithreaded filters and include many examples.
+include multithreaded filters and include many examples. Unless
+indicated otherwise, all filters will behave basically the same way as
+other kernel filters, requiring either a radius or a neighborhood to
+be specified.
 
@@
 being computed. The filters in this package are multithreaded and deal
-with arbitary dimensions. The current filters do not support the mode
+with arbitrary dimensions. The current filters do not support the mode
 of operation used in face recognition application in which an

### 2007-06-29 article/Article.tex
corrected equation refs and added ref to darcs repo
darcs-hash:20070629055044-fafb9-0559a71a743b7425fd70c7929e013a1f9bd5ec64.gz

@@
 original accumulator image. The update formulas are provided in
-Equations \ref{eq:accum}, \ref{eq:accSum},
-\ref{eq:accum2}, \ref{eq:accSum2} and \ref{eq:accumSigma}.
+Equations \ref{eq:accum2}, \ref{eq:accSum2} and \ref{eq:accumSigma}.
 
@@
 boxConvolution}. The most recent versions of these packages can be
-obtained from {\bf link to Gaetan's darcs repo}. All of the packages
+obtained from {\bf http://voxel.jouy.inra.fr/darcsweb/}. All of the packages
 include multithreaded filters and include many examples. Unless

### 2007-06-29 article/Article.tex
added timing for circle dilation
darcs-hash:20070629062554-fafb9-545b64eddf9639ce4e2dcbc36b23d67f24db12a8.gz

@@
 \subsection{Morpholgy}
+\begin{table}[htbp]
+\centering
+\begin{tabular}{cccc}
+\hline
+Radius & Histogram & vanHerk/Gil-Weman & Anchor \\
+1    &   0.0295  & 0.0228  & 0.033 \\
+2    &   0.0219  & 0.0126  & 0.0187 \\
+3    &   0.028   & 0.0128  & 0.0167 \\
+4    &   0.0339  & 0.0262  & 0.0379 \\ 
+5    &   0.0403  & 0.0264  & 0.0363 \\
+6    &   0.046   & 0.0265  & 0.038 \\
+7    &   0.052   & 0.027   & 0.0343 \\
+8    &   0.0637  & 0.0275  & 0.0344 \\
+9    &   0.0665  & 0.0387  & 0.0599 \\
+10   &   0.117   & 0.0641  & 0.142 \\
+15   &   0.103   & 0.0385  & 0.0455 \\
+20   &   0.133   & 0.0409  & 0.0445 \\
+25   &   0.164   & 0.0399  & 0.044 \\
+30   &   0.199   & 0.0403  & 0.0427 \\
+40   &   0.263   & 0.0405  & 0.0425 \\
+50   &   0.329   & 0.0403  & 0.043 \\
+100  &   0.68    & 0.0404  &  0.0472 \\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for dilation by circles using histogram, and polygon approximation based implementations applied to the $256 \times 256$ cthead image on a 1.8GHz AMD Athlon, 512KB cache.
+The polygon approximations use the default number of lines (a maximum of 6 when radius is greater than 8.\label{tab:perfPoly}}
+\end{table}
 

### 2007-06-29 article/Article.tex
a few enhancements in the articles
darcs-hash:20070629164845-2fc9d-233d0d9ff15091e9b5d9cb3d8048fc234c3903ac.gz

@@
 boxConvolution}. The most recent versions of these packages can be
-obtained from {\bf http://voxel.jouy.inra.fr/darcsweb/}. All of the packages
-include multithreaded filters and include many examples. Unless
+obtained from \url{http://voxel.jouy.inra.fr/darcsweb/}\footnote{
+The most recent versions can be obtained using darcs \cite{DarcsWebSite}
+with the command {\em darcs get http://voxel.jouy.inra.fr/\\darcs/contrib-itk/package},
+where {\em package} must be replace by {\em consolidatedMorphology},
+{\em fastRankMean} or {\em boxConvolution}.}.
+All of the packages
+include multithreaded filters and include many examples and tests. Unless
 indicated otherwise, all filters will behave basically the same way as
@@
 
+\subsection{Extended languages support}
+
+All the new filters provided are wrapped using WrapITK's external projects
+\cite{WrapITK} and have been successfully tested with python.
+
 \section{Performance}
+
+All the execution times in this section are measured on a computer with four
+Intel\textregistered Xeon\textregistered CPU cores at 2.33GHz with 4MB cache,
+running CentOS 4.4 64 bits. Unless precised in the description, the tests are
+forced to run on a single core.
+
 \subsection{Morpholgy}
@@
 
+All the kernel based filters in ITK would also benefit to hinerit a common
+super class which manage the neighborhood size and/or the kernel, and which
+provide the infrastructure for the separable operations.
+
 \section{Comments and Conclusions}

### 2007-06-29 article/Article.tex
migale citation
darcs-hash:20070629182722-2fc9d-23f0d9a912eba76ac608c8a4fb11d51fc0fc8c0b.gz
specified seems more english than precised
darcs-hash:20070629191558-2fc9d-174552a8dfe214aa4cbb7da595d12bbcad56519d.gz

@@
 Intel\textregistered Xeon\textregistered CPU cores at 2.33GHz with 4MB cache,
-running CentOS 4.4 64 bits. Unless precised in the description, the tests are
+running CentOS 4.4 64 bits. Unless specified in the description, the tests are
 forced to run on a single core.
@@
 (\url{http://mima2.jouy.inra.fr}) for providing the 3D test image.
-
+We are grateful to the INRA MIGALE bioinformatics platform
+(\url{http://migale.jouy.inra.fr}) for providing the computational resources
+used for the timing tests.
 \appendix

### 2007-07-02 article/Article.tex
add a template for the graphics
darcs-hash:20070702195110-2fc9d-680007fd7ceb06e19cb5b764d92163ce162569d6.gz

@@
 
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Dilation]{\includegraphics[scale=0.62]{consolidatedMorphology-perf2D-dilation-small}}
+\subfigure[Opening]{\includegraphics[scale=0.62]{consolidatedMorphology-perf2D-opening-small}}
+\subfigure[Gradient]{\includegraphics[scale=0.62]{consolidatedMorphology-perf2D-gradient-small}}
+\subfigure[TODO]{\includegraphics[scale=0.62]{consolidatedMorphology-perf2D-gradient-small}}
+\subfigure[TODO]{\includegraphics[scale=0.62]{consolidatedMorphology-perf2D-gradient-small}}
+\subfigure[TODO]{\includegraphics[scale=0.62]{consolidatedMorphology-perf2D-gradient-small}}
+\caption{TODO}
+\end{center}
+\end{figure}
+
 \subsection{Rank and Mean filters}

### 2007-07-02 article/Article.tex
add a few more graphics
darcs-hash:20070702214843-2fc9d-ee884ff04584fda5be296a972d63adcbfc66a12b.gz
minor grammar fixes to article
darcs-hash:20070702235032-fafb9-6a51baa1460da340327dbd7ab3353889b2fb2d30.gz

@@
 % The author and/or editor can define 'significant' however they like.
-\release{0.00}
+% \release{0.00}
 
@@
 A kernel based filtering process replaces the pixel at the kernel
-origin with the result of applying a function to all pixels defined by
+origin with the result of a function of all pixels defined by
 the kernel. Many useful filters, including edge detection and gradient
@@
 approach is very efficient when 8 or 16 bit pixels are used because
-the histogram can be represented as an array, with place holders
-tracking the current maximum or minimum increasing performance. More
-sophisticated histogram representations are necessary for larger pixel
-types. Our implementation uses c++ maps, which is an extension to the
-original paper.
+the histogram can be represented as an array and the histogram updated
+by incrementing or decrementing the appropriate bins. Using place
+holders to track the current maximum or minimum increases
+performance. More sophisticated histogram representations are
+necessary for larger pixel types. Our implementation uses c++ maps,
+which is an extension to the original paper.
 
@@
 
-Approximations of more complex shapes, particularly circles and
+Approximations of more complex shapes, notably circles and
 ellipses can constructed using a number of lines at evenly spaced
@@
 higher precision data, potentially reducing any speed advantage.
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Input]{\includegraphics[scale=1.88]{CellsFluorescence2}}
+\subfigure[Dilation]{\includegraphics[scale=0.45]{dilate2D-std}}
+\subfigure[Erosion]{\includegraphics[scale=0.45]{erode2D-std}}
+\subfigure[Opening]{\includegraphics[scale=0.45]{open2D-std}}
+\subfigure[Closing]{\includegraphics[scale=0.45]{close2D-std}}
+\subfigure[White top-hat]{\includegraphics[scale=0.45]{whiteTopHat2D-std}}
+% \subfigure[Gradient]{\includegraphics[scale=0.5]{gradient2D-std}}
+\caption{Effect of a dilation (b), an erosion (c), an opening (d),
+a closing (e) and a white top hat (f) with a ball structuring element.}
+\end{center}
+\end{figure}
 
@@
 \caption{Recursive computation of the accumulator image - value at $(i,j)$ computed using Equation \ref{eq:accum}.
-The double hashed region is subtracted twice because it is included in both $b_{i,j-1}$ and $b_{i-1, j}$.\label{fig:accum}}
+The double hashed region is subtracted because it is included in both $b_{i,j-1}$ and $b_{i-1, j}$.\label{fig:accum}}
 \end{figure}
@@
 
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Input]{\includegraphics[scale=0.6]{cthead1}}
+\subfigure[Mean]{\includegraphics[scale=0.6]{test}}
+\subfigure[Sigma]{\includegraphics[scale=0.6]{sigma}}
+\caption{A mean transform (b) and a standard deviation transform (c) with a kernel of radius 3.}
+\end{center}
+\end{figure}
+
 \section{ITK implementation}
 The filters discusses in this article are contained in 3 packages --
-{\em consolidatedMorphology}, {\em fastRankMean} and {\em
+{\em consolidatedMorphology}\footnote{An earlier version of the {\em consolidatedMorphology} package has
+been submitted to the InsightJournal. However many improvements and
+bug fixes have been included since then.}, {\em fastRankMean} and {\em
 boxConvolution}. The most recent versions of these packages can be
@@
 \subsection{Morphology filters -- consolidatedMorphology}
-An earlier version of the {\em consolidatedMorphology} package has
-been submitted to the InsightJournal. However many improvements and
-bug fixes have been included since then.
 
@@
 \begin{itemize}
-  \item \verb$Ball()$ generate a ball structuring element. It takes the 
+  \item \verb$Ball()$ generate a ball structuring element (a circle in 2D). It takes the 
 	radius of the structuring element as parameter.
@@
 method, with options of {\em BASIC}, {\em HISTO}, {\em VHGW} and {\em
-ANCHOR} for the traditional ITK implementation, the moving histogram
+ANCHOR}, for the traditional ITK implementation, the moving histogram
 algorithm, the van Herk, Gil, Werman algorithm and the anchor
@@
 used for Box and Poly structuring elements. VHGW is more thoroughly
-tested. This HISTO algorithm produces the identical results to BASIC,
+tested. The HISTO algorithm produces the identical results to BASIC,
 and is always faster. It is the default and should be used if you need
@@
 \subsection{Morpholgy}
-\begin{table}[htbp]
-\centering
-\begin{tabular}{cccc}
-\hline
-Radius & Histogram & vanHerk/Gil-Weman & Anchor \\
-1    &   0.0295  & 0.0228  & 0.033 \\
-2    &   0.0219  & 0.0126  & 0.0187 \\
-3    &   0.028   & 0.0128  & 0.0167 \\
-4    &   0.0339  & 0.0262  & 0.0379 \\ 
-5    &   0.0403  & 0.0264  & 0.0363 \\
-6    &   0.046   & 0.0265  & 0.038 \\
-7    &   0.052   & 0.027   & 0.0343 \\
-8    &   0.0637  & 0.0275  & 0.0344 \\
-9    &   0.0665  & 0.0387  & 0.0599 \\
-10   &   0.117   & 0.0641  & 0.142 \\
-15   &   0.103   & 0.0385  & 0.0455 \\
-20   &   0.133   & 0.0409  & 0.0445 \\
-25   &   0.164   & 0.0399  & 0.044 \\
-30   &   0.199   & 0.0403  & 0.0427 \\
-40   &   0.263   & 0.0405  & 0.0425 \\
-50   &   0.329   & 0.0403  & 0.043 \\
-100  &   0.68    & 0.0404  &  0.0472 \\
-\hline
-\hline
-\end{tabular}
-\caption{Execution times in seconds for dilation by circles using histogram, and polygon approximation based implementations applied to the $256 \times 256$ cthead image on a 1.8GHz AMD Athlon, 512KB cache.
-The polygon approximations use the default number of lines (a maximum of 6 when radius is greater than 8.\label{tab:perfPoly}}
-\end{table}
+
+Execution times for the dilation, the opening, and the gradient transform are shown in
+Tables \ref{tab:perfDilation}, \ref{tab:perfOpening} and \ref{tab:perfGradient}, and in
+Figure \ref{fig:consolidatedMorphology-perf}. The predicted linear complexity for the number
+of neighbors is observed for the basic algorithm, as well as the constant complexity of the
+van Herk / Gil Werman and the anchor algorithm. The linear complexity for the number of
+pixels added and removed per translation is observed for the moving histogram algorithm,
+as expected.
+
+A good threading support is observed for all the implementations, as shown in Table
+\ref{tab:perfThreads} and Figure \ref{fig:consolidatedMorphology-perf}.
 
@@
 \subfigure[Gradient]{\includegraphics[scale=0.62]{consolidatedMorphology-perf2D-gradient-small}}
-\subfigure[TODO]{\includegraphics[scale=0.62]{consolidatedMorphology-perf2D-gradient-small}}
-\subfigure[TODO]{\includegraphics[scale=0.62]{consolidatedMorphology-perf2D-gradient-small}}
-\subfigure[TODO]{\includegraphics[scale=0.62]{consolidatedMorphology-perf2D-gradient-small}}
-\caption{TODO}
+\subfigure[Threads]{\includegraphics[scale=0.62]{consolidatedMorphology-perf_threads}}
+\subfigure[Number of neighbors]{\includegraphics[scale=0.62]{consolidatedMorphology-perf_nb_of_neighbors}}
+\subfigure[Polygon]{\includegraphics[scale=0.62]{consolidatedMorphology-perf2Dpoly}}
+\caption{Execution times with a box structuring element of increasing size for different algorithms of dilation (a),
+opening (b) and gradient (c).
+Execution times with and increasing number fo threads (d).
+Execution times with an increasing number of neighbors in a box structuring of constant radius (e).
+Execution times with a polygon structuring element. The polygon approximations use the default number of lines (a maximum of 6 when radius is greater than 8 (f).\label{fig:consolidatedMorphology-perf}
+Results are reported in Tables \ref{tab:perfDilation}, \ref{tab:perfOpening}, \ref{tab:perfGradient},
+\ref{tab:perfThreads} and \ref{tab:perf2Dpoly}.
+}
 \end{center}
@@
 Relative times for sliding window median and mean filters and
-separable median filters are shown in Table \ref{tab:perfRank}. The
+separable median filters are shown in Tables \ref{tab:perfRankMean}
+and \ref{tab:perfRankMedian} and in Figure \ref{fig:perfRank}. The
 predicted linear complexity is observed for the sliding window
@@
 runtime independent of kernel size, as expected.
-\begin{table}[htbp]
-\centering
-\begin{tabular}{cccccc}
-\hline
-Size  &  Direct Median &  Sliding Median & Separable Median & Direct Mean & Sliding Mean\\
-1     &   0.0352 & 0.0341 & 0.0378 & 0.0171 & 0.0229\\
-5     &   0.409  & 0.0823  & 0.0201 & 0.24  &  0.0601 \\
-10    &   1.03   & 0.131 & 0.0193 & 0.803 &  0.0569 \\
-15    &   2.82   & 0.106  & 0.0192 & 2.13  &  0.0819 \\
-20    &   5.85   & 0.138  & 0.0358 & 4.76  &  0.106 \\
-40    &   39.7   & 0.2563  & 0.0198 & 36.6  &  0.203 \\
-\hline
-\hline
-\end{tabular}
-\caption{Execution times in seconds for median and mean filters using direct and sliding window implementations applied to the $256 \times 256$ cthead image on a 1.8GHz AMD Athlon, 512KB cache.
-The separable median is an approximation and therefore doesn't produce the same result as the direct or sliding algorithms.\label{tab:perfRank}}
-\end{table}
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Mean]{\includegraphics[scale=0.62]{fastRankMean-perfMean}}
+\subfigure[Median]{\includegraphics[scale=0.62]{fastRankMean-perfMedian}}
+\caption{Execution times for several algorithms of mean filters (a) and
+median filters (b).\label{fig:perfRank}
+Results are reported in Tables \ref{tab:perfRankMean} and  \ref{tab:perfRankMedian}.}
+\end{center}
+\end{figure}
 
@@
 The relative times for optimized and standard convolution are shown in
-Table \ref{tab:perfBoxConv}, and demonstrate significantly improved
+Table \ref{tab:perfBoxConv} and Figure \ref{fig:perfBoxConv}, and demonstrate significantly improved
 performance for kernel radii greater than 1. The times support the
@@
 independent of kernel size.
-\begin{table}[htbp]
-\centering
-\begin{tabular}{ccccc}
-\hline
-Size  &  Direct Mean &  Box Mean & Direct sigma & Box sigma\\
-1    &   0.0142 & 0.0226 & 0.0176 & 0.0332 \\
-2    &   0.0279 & 0.0176 & 0.0294 & 0.0258 \\
-3    &   0.0425 & 0.013  & 0.0447 & 0.0193 \\
-4    &   0.077  & 0.0132 & 0.0803 & 0.0195 \\
-5    &   0.128  & 0.0132 & 0.129  & 0.0195 \\
-6    &   0.195  & 0.0134 & 0.195  & 0.0198 \\
-7    &   0.275  & 0.0136 & 0.283  & 0.0198 \\
-8    &   0.392  & 0.0148 & 0.405  & 0.0204 \\
-9    &   0.54   & 0.014  & 0.548  & 0.0201 \\
-10   &   0.706  & 0.0139 & 0.713  & 0.0202 \\
-15   &   2.12   & 0.0158 & 2.18   & 0.021 \\
-20   &   4.62   & 0.0157 & 4.76   & 0.0217 \\
-25   &   8.63   & 0.0164 & 8.91   & 0.0222 \\
-\hline
-\hline
-\end{tabular}
-\caption{Execution times in seconds for mean and standard deviation using direct and box convolution implementations applied to the $256 \times 256$ cthead image on a 1.8GHz AMD Athlon, 512KB cache.\label{tab:perfBoxConv}}
-\end{table}
+
+A good threading support is observed for all the implementations, as shown in Table
+\ref{tab:perfBoxConv-threads} and Figure \ref{fig:perfBoxConv}.
+
+\begin{figure}[htbp]
+\begin{center}
+\subfigure[Neighborhood size]{\includegraphics[scale=0.62]{boxConvolution-perf2D-large}}
+\subfigure[Threads]{\includegraphics[scale=0.62]{boxConvolution-perf_threads-large}}
+\caption{Execution times for and increasing kernel size (a), and an increasing number of threads (b).
+\label{fig:perfBoxConv}
+Results are reported in Tables \ref{tab:perfBoxConv} and  \ref{tab:perfBoxConv-threads}}
+\end{center}
+\end{figure}
+
 \subsection{Timing notes}
@@
 used for the timing tests.
+
 \appendix
+
+\begin{table}[phtb]
+\centering
+\begin{tabular}{ccccccc}
+\hline
+Radius	&	Basic	&	Separable	&	Histogram	&	Separable histogram	&	Anchor	&	van Herk / Gil Werman	\\
+\hline
+1	&	0.00477	&	0.00491	&	0.013	&	0.012	&	0.0144	&	0.00725	\\
+2	&	0.0114	&	0.0101	&	0.0161	&	0.0119	&	0.0135	&	0.00743	\\
+3	&	0.024	&	0.0146	&	0.0193	&	0.0116	&	0.0129	&	0.00794	\\
+4	&	0.0444	&	0.0192	&	0.0231	&	0.0114	&	0.0126	&	0.00755	\\
+5	&	0.0714	&	0.0247	&	0.0261	&	0.0114	&	0.0121	&	0.0074	\\
+6	&	0.107	&	0.0299	&	0.0313	&	0.0116	&	0.0127	&	0.00771	\\
+7	&	0.153	&	0.034	&	0.0358	&	0.0113	&	0.0121	&	0.00859	\\
+8	&	0.211	&	0.0407	&	0.038	&	0.0112	&	0.0121	&	0.00773	\\
+9	&	0.281	&	0.0442	&	0.0434	&	0.0114	&	0.012	&	0.00779	\\
+10	&	0.365	&	0.0486	&	0.0446	&	0.0111	&	0.0118	&	0.00763	\\
+15	&	1.03	&	0.074	&	0.0604	&	0.0115	&	0.0116	&	0.00794	\\
+20	&	2.23	&	0.104	&	0.079	&	0.0113	&	0.0113	&	0.00762	\\
+25	&	4.11	&	0.129	&	0.0968	&	0.0114	&	0.0116	&	0.0079	\\
+30	&	6.82	&	0.154	&	0.111	&	0.0114	&	0.0115	&	0.00781	\\
+40	&	15.4	&	0.214	&	0.153	&	0.0116	&	0.0116	&	0.00785	\\
+50	&	29.3	&	0.285	&	0.19	&	0.0119	&	0.0117	&	0.00804	\\
+100	&	226	&	0.767	&	0.355	&	0.0129	&	0.0117	&	0.00809	\\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for dilation by boxes using several algorithm implementations applied to the $256 \times 256$ cthead image.\label{tab:perfDilation}}
+\end{table}
+
+\begin{table}[phtb]
+\centering
+\begin{tabular}{ccccc}
+\hline
+Radius	&	Basic	&	Histogram	&	Anchor	&	van Herk / Gil Werman	\\
+\hline
+1	&	0.0107	&	0.0124	&	0.0219	&	0.014	\\
+2	&	0.0235	&	0.0167	&	0.0201	&	0.0144	\\
+3	&	0.0488	&	0.0211	&	0.019	&	0.0154	\\
+4	&	0.0895	&	0.0247	&	0.0184	&	0.0148	\\
+5	&	0.143	&	0.0273	&	0.0176	&	0.0143	\\
+6	&	0.216	&	0.037	&	0.0184	&	0.0149	\\
+7	&	0.308	&	0.0372	&	0.0176	&	0.0167	\\
+8	&	0.423	&	0.0401	&	0.0172	&	0.015	\\
+9	&	0.563	&	0.0466	&	0.0171	&	0.0151	\\
+10	&	0.73	&	0.0455	&	0.0166	&	0.0148	\\
+15	&	2.07	&	0.0624	&	0.0165	&	0.0153	\\
+20	&	4.46	&	0.0885	&	0.0159	&	0.0148	\\
+25	&	8.22	&	0.104	&	0.0163	&	0.0153	\\
+30	&	13.6	&	0.118	&	0.0161	&	0.0151	\\
+40	&	30.9	&	0.153	&	0.0163	&	0.0153	\\
+50	&	58.6	&	0.196	&	0.0168	&	0.0156	\\
+100	&	452	&	0.339	&	0.0163	&	0.0157	\\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for morphological gradient transformation by boxes using several algorithm implementations applied to the $256 \times 256$ cthead image.\label{tab:perfGradient}}
+\end{table}
+
+\begin{table}[phtb]
+\centering
+\begin{tabular}{ccccc}
+\hline
+Radius	&	Basic	&	Histogram	&	Anchor	&	van Herk / Gil Werman	\\
+\hline
+1	&	0.0105	&	0.0208	&	0.0167	&	0.0146	\\
+2	&	0.0217	&	0.0283	&	0.0157	&	0.0148	\\
+3	&	0.0424	&	0.0352	&	0.0145	&	0.0154	\\
+4	&	0.0752	&	0.0444	&	0.0142	&	0.0159	\\
+5	&	0.117	&	0.049	&	0.0143	&	0.0159	\\
+6	&	0.172	&	0.0573	&	0.014	&	0.0162	\\
+7	&	0.24	&	0.0676	&	0.0139	&	0.0165	\\
+8	&	0.324	&	0.072	&	0.0141	&	0.0169	\\
+9	&	0.425	&	0.0804	&	0.0141	&	0.0175	\\
+10	&	0.546	&	0.093	&	0.0143	&	0.0176	\\
+15	&	1.5	&	0.14	&	0.0154	&	0.019	\\
+20	&	3.19	&	0.178	&	0.0155	&	0.0204	\\
+25	&	5.9	&	0.235	&	0.0164	&	0.0219	\\
+30	&	9.9	&	0.274	&	0.0175	&	0.0236	\\
+40	&	23	&	0.383	&	0.0193	&	0.0272	\\
+50	&	45.3	&	0.522	&	0.0213	&	0.0298	\\
+100	&	421	&	1.6	&	0.0346&		0.0493	\\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for morphological opening by boxes using several algorithm implementations applied to the $256 \times 256$ cthead image.\label{tab:perfOpening}}
+\end{table}
+
+\begin{table}[phtb]
+\centering
+\begin{tabular}{ccccc}
+\hline
+Threads	&	Basic	&	Histogram	&	Anchor	&	van Herk / Gil Werman	\\
+\hline
+1	&	3.37	&	2.34	&	1.23	&	1.02	\\
+2	&	1.69	&	1.64	&	0.73	&	0.615	\\
+3	&	1.14	&	1.15	&	0.46	&	0.367	\\
+4	&	0.871	&	0.866	&	0.39	&	0.312	\\
+5	&	1.06	&	1.03	&	0.505	&	0.437	\\
+6	&	1.01	&	0.949	&	0.489	&	0.422	\\
+7	&	0.963	&	0.866	&	0.493	&	0.422	\\
+8	&	0.918	&	0.849	&	0.506	&	0.453	\\
+9	&	0.947	&	0.897	&	0.548	&	0.491	\\
+10	&	0.92	&	0.874	&	0.568	&	0.506	\\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for morphological opening by boxes using several algorithm implementations applied to the $3200 \times 2400$ image.\label{tab:perfThreads}}
+\end{table}
+
+\begin{table}[phtb]
+\centering
+\begin{tabular}{cccc}
+\hline
+Radius	&	Histogram	&	van Herk / Gil Werma	&	Anchorn	\\
+\hline
+1	&	0.354	&	0.379	&	0.532	\\
+2	&	0.486	&	0.329	&	0.455	\\
+3	&	0.62	&	0.338	&	0.411	\\
+4	&	0.74	&	0.763	&	0.995	\\
+5	&	0.862	&	0.757	&	0.95	\\
+6	&	1.03	&	0.759	&	0.968	\\
+7	&	1.17	&	0.77	&	0.936	\\
+8	&	1.31	&	0.768	&	0.894	\\
+9	&	1.52	&	1.11	&	1.28	\\
+10	&	1.67	&	1.08	&	1.31	\\
+15	&	2.42	&	1.09	&	1.25	\\
+20	&	3.13	&	1.16	&	1.29	\\
+25	&	3.84	&	1.14	&	1.26	\\
+30	&	4.64	&	1.17	&	1.3	\\
+40	&	5.95	&	1.19	&	1.23	\\
+50	&	7.6	&	1.22	&	1.41	\\
+100	&	19	&	1.36	&	1.34	\\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for morphological opening by boxes using several algorithm implementations applied to the $256 \times 256$ cthead image.\label{tab:perf2Dpoly}}
+\end{table}
+
+\begin{table}[phtb]
+\centering
+\begin{tabular}{ccccc}
+\hline
+Radius	&	Direct	&	Moving window	&	Separable direct	&	Separable moving window	\\
+\hline
+1	&	0.00355	&	0.00498	&	0.00487	&	0.00668	\\
+2	&	0.00642	&	0.00706	&	0.0054	&	0.00679	\\
+3	&	0.0127	&	0.00761	&	0.00861	&	0.00756	\\
+5	&	0.0259	&	0.00936	&	0.00803	&	0.00749	\\
+10	&	0.121	&	0.0162	&	0.0129	&	0.00765	\\
+15	&	0.35	&	0.0224	&	0.0206	&	0.00755	\\
+20	&	0.771	&	0.0252	&	0.0278	&	0.0076	\\
+40	&	5.59	&	0.059	&	0.0699	&	0.00871	\\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for mean filters using direct and sliding window implementations applied to the $256 \times 256$ cthead image.\label{tab:perfRankMean}}
+\end{table}
+
+\begin{table}[phtb]
+\centering
+\begin{tabular}{ccccc}
+\hline
+Radius	&	Direct	&	Huang	&	Separable direct	&	Separable huang	\\
+\hline
+1	&	0.00642	&	0.00758	&	0.0071	&	0.0093	\\
+5	&	0.0448	&	0.0186	&	0.0153	&	0.00913	\\
+10	&	0.177	&	0.0265	&	0.0291	&	0.00895	\\
+15	&	0.445	&	0.0342	&	0.0362	&	0.00876	\\
+20	&	0.93	&	0.043	&	0.0444	&	0.00925	\\
+40	&	6.21	&	0.061	&	0.0899	&	0.00936	\\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for median filters using direct and sliding window implementations applied to the $256 \times 256$ cthead image.
+The separable median is an approximation and therefore doesn't produce the same result as the direct or sliding algorithms.\label{tab:perfRankMedian}}
+\end{table}
+
+\begin{table}[phtb]
+\centering
+\begin{tabular}{ccccc}
+\hline
+Size  &  Direct Mean &  Box Mean & Direct sigma & Box sigma\\
+1	&	0.557	&	0.916	&	0.858	&	1.45	 \\
+2	&	1.15	&	0.924	&	1.64	&	1.56	 \\
+3	&	2.08	&	0.932	&	2.74	&	1.56	 \\
+4	&	3.47	&	0.92	&	4.34	&	1.55	 \\
+5	&	5.13	&	0.935	&	6.34	&	1.55	 \\
+6	&	7.06	&	1.08	&	8.53	&	1.62	 \\
+7	&	9.42	&	1.08	&	11.3	&	1.63	 \\
+8	&	12.2	&	0.922	&	14.5	&	1.63	 \\
+9	&	15.3	&	0.924	&	18	&	1.63	 \\
+10	&	18.9	&	0.925	&	22.1	&	1.63	 \\
+15	&	43.3	&	0.929	&	50.2	&	1.64	 \\
+20	&	80.1	&	0.936	&	91.8	&	1.64	 \\
+25	&	131	&	0.939	&	149	&	1.64	 \\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for mean and standard deviation using direct and box convolution implementations applied to the $3200 \times 2400$ image.\label{tab:perfBoxConv}}
+\end{table}
+
+\begin{table}[phtb]
+\centering
+\begin{tabular}{ccccc}
+\hline
+Size  &  Direct Mean &  Box Mean & Direct sigma & Box sigma\\
+\hline
+1	&	2.1	&	0.942	&	2.75	&	1.57	\\
+2	&	1.06	&	0.466	&	1.38	&	0.805	\\
+3	&	0.716	&	0.391	&	0.932	&	0.551	\\
+4	&	0.538	&	0.335	&	0.846	&	0.453	\\
+5	&	0.659	&	0.327	&	0.902	&	0.525	\\
+6	&	0.616	&	0.298	&	0.728	&	0.47	\\
+7	&	0.579	&	0.285	&	0.758	&	0.466	\\
+8	&	0.558	&	0.27	&	0.729	&	0.473	\\
+9	&	0.614	&	0.289	&	0.771	&	0.467	\\
+10	&	0.578	&	0.281	&	0.732	&	0.476	\\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for mean and standard deviation using direct and box convolution implementations applied to the $3200 \times 2400$ image.\label{tab:perfBoxConv-threads}}
+\end{table}
 

### 2007-07-03 article/Article.tex
timing results continued
darcs-hash:20070703133114-2fc9d-bf2f64a2ed4148139bc26446ef0d1eeab89a3a4e.gz

@@
 % The author and/or editor can define 'significant' however they like.
-% \release{0.00}
+\release{0.00}
 
@@
 A kernel based filtering process replaces the pixel at the kernel
-origin with the result of a function of all pixels defined by
+origin with the result of applying a function to all pixels defined by
 the kernel. Many useful filters, including edge detection and gradient
@@
 approach is very efficient when 8 or 16 bit pixels are used because
-the histogram can be represented as an array and the histogram updated
-by incrementing or decrementing the appropriate bins. Using place
-holders to track the current maximum or minimum increases
-performance. More sophisticated histogram representations are
-necessary for larger pixel types. Our implementation uses c++ maps,
-which is an extension to the original paper.
+the histogram can be represented as an array, with place holders
+tracking the current maximum or minimum increasing performance. More
+sophisticated histogram representations are necessary for larger pixel
+types. Our implementation uses c++ maps, which is an extension to the
+original paper.
 
@@
 
-Approximations of more complex shapes, notably circles and
+Approximations of more complex shapes, particularly circles and
 ellipses can constructed using a number of lines at evenly spaced
@@
 higher precision data, potentially reducing any speed advantage.
-
-\begin{figure}[htbp]
-\begin{center}
-\subfigure[Input]{\includegraphics[scale=1.88]{CellsFluorescence2}}
-\subfigure[Dilation]{\includegraphics[scale=0.45]{dilate2D-std}}
-\subfigure[Erosion]{\includegraphics[scale=0.45]{erode2D-std}}
-\subfigure[Opening]{\includegraphics[scale=0.45]{open2D-std}}
-\subfigure[Closing]{\includegraphics[scale=0.45]{close2D-std}}
-\subfigure[White top-hat]{\includegraphics[scale=0.45]{whiteTopHat2D-std}}
-% \subfigure[Gradient]{\includegraphics[scale=0.5]{gradient2D-std}}
-\caption{Effect of a dilation (b), an erosion (c), an opening (d),
-a closing (e) and a white top hat (f) with a ball structuring element.}
-\end{center}
-\end{figure}
 
@@
 \caption{Recursive computation of the accumulator image - value at $(i,j)$ computed using Equation \ref{eq:accum}.
-The double hashed region is subtracted because it is included in both $b_{i,j-1}$ and $b_{i-1, j}$.\label{fig:accum}}
+The double hashed region is subtracted twice because it is included in both $b_{i,j-1}$ and $b_{i-1, j}$.\label{fig:accum}}
 \end{figure}
@@
 
-\begin{figure}[htbp]
-\begin{center}
-\subfigure[Input]{\includegraphics[scale=0.6]{cthead1}}
-\subfigure[Mean]{\includegraphics[scale=0.6]{test}}
-\subfigure[Sigma]{\includegraphics[scale=0.6]{sigma}}
-\caption{A mean transform (b) and a standard deviation transform (c) with a kernel of radius 3.}
-\end{center}
-\end{figure}
-
 \section{ITK implementation}
 The filters discusses in this article are contained in 3 packages --
-{\em consolidatedMorphology}\footnote{An earlier version of the {\em consolidatedMorphology} package has
-been submitted to the InsightJournal. However many improvements and
-bug fixes have been included since then.}, {\em fastRankMean} and {\em
+{\em consolidatedMorphology}, {\em fastRankMean} and {\em
 boxConvolution}. The most recent versions of these packages can be
@@
 \subsection{Morphology filters -- consolidatedMorphology}
+An earlier version of the {\em consolidatedMorphology} package has
+been submitted to the InsightJournal. However many improvements and
+bug fixes have been included since then.
 
@@
 \begin{itemize}
-  \item \verb$Ball()$ generate a ball structuring element (a circle in 2D). It takes the 
+  \item \verb$Ball()$ generate a ball structuring element. It takes the 
 	radius of the structuring element as parameter.
@@
 method, with options of {\em BASIC}, {\em HISTO}, {\em VHGW} and {\em
-ANCHOR}, for the traditional ITK implementation, the moving histogram
+ANCHOR} for the traditional ITK implementation, the moving histogram
 algorithm, the van Herk, Gil, Werman algorithm and the anchor
@@
 used for Box and Poly structuring elements. VHGW is more thoroughly
-tested. The HISTO algorithm produces the identical results to BASIC,
+tested. This HISTO algorithm produces the identical results to BASIC,
 and is always faster. It is the default and should be used if you need
@@
 Relative times for sliding window median and mean filters and
-separable median filters are shown in Tables \ref{tab:perfRankMean}
-and \ref{tab:perfRankMedian} and in Figure \ref{fig:perfRank}. The
+separable median filters are shown in Table \ref{tab:perfRank}. The
 predicted linear complexity is observed for the sliding window
@@
 \subfigure[Median]{\includegraphics[scale=0.62]{fastRankMean-perfMedian}}
-\caption{Execution times for several algorithms of mean filters (a) and
-median filters (b).\label{fig:perfRank}
-Results are reported in Tables \ref{tab:perfRankMean} and  \ref{tab:perfRankMedian}.}
+\caption{TODO}
 \end{center}
@@
 \label{fig:perfBoxConv}
-Results are reported in Tables \ref{tab:perfBoxConv} and  \ref{tab:perfBoxConv-threads}}
+Results are reported in Table \ref{tab:perfBoxConv}}
 \end{center}
@@
 \hline
-Radius	&	Direct	&	Moving window	&	Separable direct	&	Separable moving window	\\
-\hline
-1	&	0.00355	&	0.00498	&	0.00487	&	0.00668	\\
-2	&	0.00642	&	0.00706	&	0.0054	&	0.00679	\\
-3	&	0.0127	&	0.00761	&	0.00861	&	0.00756	\\
-5	&	0.0259	&	0.00936	&	0.00803	&	0.00749	\\
-10	&	0.121	&	0.0162	&	0.0129	&	0.00765	\\
-15	&	0.35	&	0.0224	&	0.0206	&	0.00755	\\
-20	&	0.771	&	0.0252	&	0.0278	&	0.0076	\\
-40	&	5.59	&	0.059	&	0.0699	&	0.00871	\\
-\hline
-\hline
-\end{tabular}
-\caption{Execution times in seconds for mean filters using direct and sliding window implementations applied to the $256 \times 256$ cthead image.\label{tab:perfRankMean}}
-\end{table}
-
-\begin{table}[phtb]
-\centering
-\begin{tabular}{ccccc}
-\hline
-Radius	&	Direct	&	Huang	&	Separable direct	&	Separable huang	\\
-\hline
-1	&	0.00642	&	0.00758	&	0.0071	&	0.0093	\\
-5	&	0.0448	&	0.0186	&	0.0153	&	0.00913	\\
-10	&	0.177	&	0.0265	&	0.0291	&	0.00895	\\
-15	&	0.445	&	0.0342	&	0.0362	&	0.00876	\\
-20	&	0.93	&	0.043	&	0.0444	&	0.00925	\\
-40	&	6.21	&	0.061	&	0.0899	&	0.00936	\\
-\hline
-\hline
-\end{tabular}
-\caption{Execution times in seconds for median filters using direct and sliding window implementations applied to the $256 \times 256$ cthead image.
-The separable median is an approximation and therefore doesn't produce the same result as the direct or sliding algorithms.\label{tab:perfRankMedian}}
-\end{table}
-
-\begin{table}[phtb]
-\centering
-\begin{tabular}{ccccc}
-\hline
 Size  &  Direct Mean &  Box Mean & Direct sigma & Box sigma\\
@@
 
+\begin{table}[phtb]
+\centering
+\begin{tabular}{ccccc}
+\hline
+Radius	&	Direct	&	Moving window	&	Separable direct	&	Separable moving window	\\
+\hline
+1	&	0.00355	&	0.00498	&	0.00487	&	0.00668	\\
+2	&	0.00642	&	0.00706	&	0.0054	&	0.00679	\\
+3	&	0.0127	&	0.00761	&	0.00861	&	0.00756	\\
+5	&	0.0259	&	0.00936	&	0.00803	&	0.00749	\\
+10	&	0.121	&	0.0162	&	0.0129	&	0.00765	\\
+15	&	0.35	&	0.0224	&	0.0206	&	0.00755	\\
+20	&	0.771	&	0.0252	&	0.0278	&	0.0076	\\
+40	&	5.59	&	0.059	&	0.0699	&	0.00871	\\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for mean filters using direct and sliding window implementations applied to the $256 \times 256$ cthead image.\label{tab:perfRankMean}}
+\end{table}
+
+\begin{table}[phtb]
+\centering
+\begin{tabular}{ccccc}
+\hline
+Radius	&	Direct	&	Huang	&	Separable direct	&	Separable huang	\\
+\hline
+1	&	0.00642	&	0.00758	&	0.0071	&	0.0093	\\
+5	&	0.0448	&	0.0186	&	0.0153	&	0.00913	\\
+10	&	0.177	&	0.0265	&	0.0291	&	0.00895	\\
+15	&	0.445	&	0.0342	&	0.0362	&	0.00876	\\
+20	&	0.93	&	0.043	&	0.0444	&	0.00925	\\
+40	&	6.21	&	0.061	&	0.0899	&	0.00936	\\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for median filters using direct and sliding window implementations applied to the $256 \times 256$ cthead image.
+The separable median is an approximation and therefore doesn't produce the same result as the direct or sliding algorithms.\label{tab:perfRankMedian}}
+\end{table}
+
 \bibliographystyle{plain}

### 2007-07-03 article/Article.tex
add a few images, to illustrate the paper
darcs-hash:20070703140234-2fc9d-ba80dd3cb0c7125057a6f4d80f77232b89222323.gz

--- 2007-07-02 article/Article.tex
@@
 \caption{Execution times for several algorithms of mean filters (a) and
-median filters (b).\label{fig:perfRank}
-Results are reported in Tables \ref{tab:perfRankMean} and  \ref{tab:perfRankMedian}.}
+median filters (b).\label{fig:perfRank}}
 \end{center}

### 2007-07-03 article/Article.tex
missing table reference
darcs-hash:20070703140948-2fc9d-5b0237a5b52710a1fe406aad0b5b35424ad73650.gz
missing table reference
darcs-hash:20070703141204-2fc9d-70b1992fea85c74c46cdb1593a919e94f2d3678a.gz

@@
 \caption{Execution times for several algorithms of mean filters (a) and
-median filters (b).\label{fig:perfRank}}
+median filters (b).\label{fig:perfRank}
+Results are reported in Tables \ref{tab:perfRankMean} and  \ref{tab:perfRankMedian}.}
 \end{center}
@@
 \label{fig:perfBoxConv}
-Results are reported in Table \ref{tab:perfBoxConv}}
+Results are reported in Tables \ref{tab:perfBoxConv} and  \ref{tab:perfBoxConv-threads}}
 \end{center}
@@
 \hline
+Radius	&	Direct	&	Moving window	&	Separable direct	&	Separable moving window	\\
+\hline
+1	&	0.00355	&	0.00498	&	0.00487	&	0.00668	\\
+2	&	0.00642	&	0.00706	&	0.0054	&	0.00679	\\
+3	&	0.0127	&	0.00761	&	0.00861	&	0.00756	\\
+5	&	0.0259	&	0.00936	&	0.00803	&	0.00749	\\
+10	&	0.121	&	0.0162	&	0.0129	&	0.00765	\\
+15	&	0.35	&	0.0224	&	0.0206	&	0.00755	\\
+20	&	0.771	&	0.0252	&	0.0278	&	0.0076	\\
+40	&	5.59	&	0.059	&	0.0699	&	0.00871	\\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for mean filters using direct and sliding window implementations applied to the $256 \times 256$ cthead image.\label{tab:perfRankMean}}
+\end{table}
+
+\begin{table}[phtb]
+\centering
+\begin{tabular}{ccccc}
+\hline
+Radius	&	Direct	&	Huang	&	Separable direct	&	Separable huang	\\
+\hline
+1	&	0.00642	&	0.00758	&	0.0071	&	0.0093	\\
+5	&	0.0448	&	0.0186	&	0.0153	&	0.00913	\\
+10	&	0.177	&	0.0265	&	0.0291	&	0.00895	\\
+15	&	0.445	&	0.0342	&	0.0362	&	0.00876	\\
+20	&	0.93	&	0.043	&	0.0444	&	0.00925	\\
+40	&	6.21	&	0.061	&	0.0899	&	0.00936	\\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for median filters using direct and sliding window implementations applied to the $256 \times 256$ cthead image.
+The separable median is an approximation and therefore doesn't produce the same result as the direct or sliding algorithms.\label{tab:perfRankMedian}}
+\end{table}
+
+\begin{table}[phtb]
+\centering
+\begin{tabular}{ccccc}
+\hline
 Size  &  Direct Mean &  Box Mean & Direct sigma & Box sigma\\
@@
 
-\begin{table}[phtb]
-\centering
-\begin{tabular}{ccccc}
-\hline
-Radius	&	Direct	&	Moving window	&	Separable direct	&	Separable moving window	\\
-\hline
-1	&	0.00355	&	0.00498	&	0.00487	&	0.00668	\\
-2	&	0.00642	&	0.00706	&	0.0054	&	0.00679	\\
-3	&	0.0127	&	0.00761	&	0.00861	&	0.00756	\\
-5	&	0.0259	&	0.00936	&	0.00803	&	0.00749	\\
-10	&	0.121	&	0.0162	&	0.0129	&	0.00765	\\
-15	&	0.35	&	0.0224	&	0.0206	&	0.00755	\\
-20	&	0.771	&	0.0252	&	0.0278	&	0.0076	\\
-40	&	5.59	&	0.059	&	0.0699	&	0.00871	\\
-\hline
-\hline
-\end{tabular}
-\caption{Execution times in seconds for mean filters using direct and sliding window implementations applied to the $256 \times 256$ cthead image.\label{tab:perfRankMean}}
-\end{table}
-
-\begin{table}[phtb]
-\centering
-\begin{tabular}{ccccc}
-\hline
-Radius	&	Direct	&	Huang	&	Separable direct	&	Separable huang	\\
-\hline
-1	&	0.00642	&	0.00758	&	0.0071	&	0.0093	\\
-5	&	0.0448	&	0.0186	&	0.0153	&	0.00913	\\
-10	&	0.177	&	0.0265	&	0.0291	&	0.00895	\\
-15	&	0.445	&	0.0342	&	0.0362	&	0.00876	\\
-20	&	0.93	&	0.043	&	0.0444	&	0.00925	\\
-40	&	6.21	&	0.061	&	0.0899	&	0.00936	\\
-\hline
-\hline
-\end{tabular}
-\caption{Execution times in seconds for median filters using direct and sliding window implementations applied to the $256 \times 256$ cthead image.
-The separable median is an approximation and therefore doesn't produce the same result as the direct or sliding algorithms.\label{tab:perfRankMedian}}
-\end{table}
-
 \bibliographystyle{plain}

### 2007-07-04 article/Article.tex
moved appendix to end and added title
darcs-hash:20070704063514-fafb9-445848776be88cfb82ca616a723e9b5418ef257f.gz

@@
 
+\bibliographystyle{plain}
+\bibliography{InsightJournal,local}
+\nocite{ITKSoftwareGuide}
+
 \appendix
+\section*{Appendix}
 
@@
 
-\bibliographystyle{plain}
-\bibliography{InsightJournal,local}
-\nocite{ITKSoftwareGuide}
-
 \end{document}

### 2007-08-01 itkBoxAccumulatorImageFilter.h
use the BoxImageFilter as superclass
darcs-hash:20070801094123-2fc9d-f44522dde6a62be8f37b557b00c713bd51426c19.gz

@@ -51,8 +51,6 @@ public:
   /** Image related typedefs. */
   itkStaticConstMacro(ImageDimension, unsigned int,
                       TInputImage::ImageDimension);
-  /** n-dimensional Kernel radius. */
-  typedef typename TInputImage::SizeType RadiusType ;
 
   void GenerateData();
 
@@ -60,13 +58,9 @@ protected:
   BoxAccumulatorImageFilter();
   ~BoxAccumulatorImageFilter() {};
 
-  void PrintSelf(std::ostream& os, Indent indent) const;
-
 private:
   BoxAccumulatorImageFilter(const Self&); //purposely not implemented
   void operator=(const Self&); //purposely not implemented
-
-  RadiusType m_Radius;
 
 };
 

### 2007-08-01 itkBoxImageFilter.h

@@ -1,0 +1,84 @@
+#ifndef __itkBoxImageFilter_h
+#define __itkBoxImageFilter_h
+
+#include "itkImageToImageFilter.h"
+#include "itkCastImageFilter.h"
+
+
+namespace itk {
+
+/**
+ * \class BoxImageFilter
+ * \brief A base class for all the filters working on a box neighborhood
+ *
+ * This filter provides the code to store the radius information about the
+ * neighborhood used in the subclasses.
+ * It also conveniently reimplement the GenerateInputRequestedRegion() so
+ * that region is well defined for the provided radius.
+ *
+ * \author Gaetan Lehmann
+ */
+
+template<class TInputImage, class TOutputImage>
+class ITK_EXPORT BoxImageFilter : 
+public ImageToImageFilter<TInputImage, TOutputImage>
+{
+public:
+  /** Standard class typedefs. */
+  typedef BoxImageFilter Self;
+  typedef ImageToImageFilter<TInputImage,TOutputImage>  Superclass;
+  typedef SmartPointer<Self>        Pointer;
+  typedef SmartPointer<const Self>  ConstPointer;
+  
+  /** Standard New method. */
+  itkNewMacro(Self);
+
+  /** Runtime information support. */
+  itkTypeMacro(BoxImageFilter,
+               ImageToImageFilter);
+ 
+  /** Image related typedefs. */
+  typedef TInputImage InputImageType;
+  typedef typename TInputImage::RegionType RegionType ;
+  typedef typename TInputImage::SizeType SizeType ;
+  typedef typename TInputImage::IndexType IndexType ;
+  typedef typename TInputImage::OffsetType OffsetType ;
+
+  typedef typename TInputImage::PixelType InputPixelType ;
+  
+  typedef TOutputImage OutputImageType;
+  typedef typename TOutputImage::PixelType OutputPixelType ;
+
+  /** Image related typedefs. */
+  itkStaticConstMacro(ImageDimension, unsigned int,
+                      TInputImage::ImageDimension);
+  /** n-dimensional Kernel radius. */
+  typedef typename TInputImage::SizeType RadiusType ;
+
+  virtual void SetRadius( const RadiusType & radius );
+  virtual void SetRadius( const unsigned long & radius );
+  itkGetConstReferenceMacro(Radius, RadiusType);
+
+  void GenerateInputRequestedRegion() ;
+
+protected:
[…]

### 2007-08-01 itkBoxMeanCalculatorImageFilter.h

@@ -1,7 +1,7 @@
 #ifndef __itkBoxMeanCalculatorImageFilter_h
 #define __itkBoxMeanCalculatorImageFilter_h
 
-#include "itkImageToImageFilter.h"
+#include "itkBoxImageFilter.h"
 
 namespace itk {
 
@@ -14,12 +14,12 @@ namespace itk {
 
 template<class TInputImage, class TOutputImage=TInputImage>
 class ITK_EXPORT BoxMeanCalculatorImageFilter : 
-    public ImageToImageFilter<TInputImage, TOutputImage>
+    public BoxImageFilter<TInputImage, TOutputImage>
 {
 public:
   /** Standard class typedefs. */
   typedef BoxMeanCalculatorImageFilter Self;
-  typedef ImageToImageFilter<TInputImage, TOutputImage>  Superclass;
+  typedef BoxImageFilter<TInputImage, TOutputImage>  Superclass;
   typedef SmartPointer<Self>        Pointer;
   typedef SmartPointer<const Self>  ConstPointer;
   
@@ -28,7 +28,7 @@ public:
 
   /** Runtime information support. */
   itkTypeMacro(BoxMeanCalculatorImageFilter, 
-               BoxMeanCalculatorImageFilter);
+               BoxImageFilter);
   
   /** Image related typedefs. */
   typedef TInputImage InputImageType;
@@ -60,14 +60,7 @@ public:
 #endif
 
                         
-  /** n-dimensional Kernel radius. */
-  typedef typename TInputImage::SizeType RadiusType ;
-  itkSetMacro(Radius, RadiusType)
-  itkGetConstReferenceMacro(Radius, RadiusType)
-
-
 protected:
-  RadiusType m_Radius;
   BoxMeanCalculatorImageFilter();
   ~BoxMeanCalculatorImageFilter() {};
 

### 2007-08-01 itkBoxMeanImageFilter.h

@@ -1,7 +1,7 @@
 #ifndef __itkBoxMeanImageFilter_h
 #define __itkBoxMeanImageFilter_h
 
-#include "itkImageToImageFilter.h"
+#include "itkBoxImageFilter.h"
 
 namespace itk {
 
@@ -14,21 +14,21 @@ namespace itk {
 
 template<class TInputImage, class TOutputImage=TInputImage>
 class ITK_EXPORT BoxMeanImageFilter : 
-    public ImageToImageFilter<TInputImage, TOutputImage>
+    public BoxImageFilter<TInputImage, TOutputImage>
 {
 public:
   /** Standard class typedefs. */
   typedef BoxMeanImageFilter Self;
-  typedef ImageToImageFilter<TInputImage, TOutputImage>  Superclass;
+  typedef BoxImageFilter<TInputImage, TOutputImage>  Superclass;
   typedef SmartPointer<Self>        Pointer;
   typedef SmartPointer<const Self>  ConstPointer;
   
   /** Standard New method. */
-  itkNewMacro(Self);  
+  itkNewMacro(Self);
 
   /** Runtime information support. */
   itkTypeMacro(BoxMeanImageFilter, 
-               BoxMeanImageFilter);
+               BoxImageFilter);
   
   /** Image related typedefs. */
   typedef TInputImage InputImageType;
@@ -58,23 +58,7 @@ public:
 #endif
 
     
-  /** n-dimensional Kernel radius. in voxels */
-  typedef typename TInputImage::SizeType RadiusType ;
-  void SetRadius(unsigned R)
-  {
-    m_Radius.Fill(R);
-  }
-  
-
-
-  itkSetMacro(Radius, RadiusType)
-  itkGetConstReferenceMacro(Radius, RadiusType)
-
-  void GenerateInputRequestedRegion() ;
-    
-
 protected:
-  RadiusType m_Radius;
   BoxMeanImageFilter();
   ~BoxMeanImageFilter() {};
 

### 2007-08-01 itkBoxSigmaImageFilter.h

@@ -1,7 +1,7 @@
 #ifndef __itkBoxSigmaImageFilter_h
 #define __itkBoxSigmaImageFilter_h
 
-#include "itkImageToImageFilter.h"
+#include "itkBoxImageFilter.h"
 
 namespace itk {
 
@@ -14,12 +14,12 @@ namespace itk {
 
 template<class TInputImage, class TOutputImage=TInputImage>
 class ITK_EXPORT BoxSigmaImageFilter : 
-    public ImageToImageFilter<TInputImage, TOutputImage>
+    public BoxImageFilter<TInputImage, TOutputImage>
 {
 public:
   /** Standard class typedefs. */
   typedef BoxSigmaImageFilter Self;
-  typedef ImageToImageFilter<TInputImage, TOutputImage>  Superclass;
+  typedef BoxImageFilter<TInputImage, TOutputImage>  Superclass;
   typedef SmartPointer<Self>        Pointer;
   typedef SmartPointer<const Self>  ConstPointer;
   
@@ -28,7 +28,7 @@ public:
 
   /** Runtime information support. */
   itkTypeMacro(BoxSigmaImageFilter, 
-               BoxSigmaImageFilter);
+               BoxImageFilter);
   
   /** Image related typedefs. */
   typedef TInputImage InputImageType;
@@ -58,23 +58,7 @@ public:
 #endif
 
     
-  /** n-dimensional Kernel radius. in voxels */
-  typedef typename TInputImage::SizeType RadiusType ;
-  void SetRadius(unsigned R)
-  {
-    m_Radius.Fill(R);
-  }
-  
-
-
-  itkSetMacro(Radius, RadiusType)
-  itkGetConstReferenceMacro(Radius, RadiusType)
-
-  void GenerateInputRequestedRegion() ;
-    
-
 protected:
-  RadiusType m_Radius;
   BoxSigmaImageFilter();
   ~BoxSigmaImageFilter() {};
 

### 2007-08-01 itkMeanImageFilter.h
add the mean filter modified to use BoxImageFilter as superclass
darcs-hash:20070801122114-2fc9d-4977ed71c744834de0cff55c3d7b32f56358f802.gz

@@ -1,0 +1,113 @@
+/*=========================================================================
+
+  Program:   Insight Segmentation & Registration Toolkit
+  Module:    $RCSfile: itkMeanImageFilter.h,v $
+  Language:  C++
+  Date:      $Date: 2006/03/29 14:53:40 $
+  Version:   $Revision: 1.5 $
+
+  Copyright (c) Insight Software Consortium. All rights reserved.
+  See ITKCopyright.txt or http://www.itk.org/HTML/Copyright.htm for details.
+
+     This software is distributed WITHOUT ANY WARRANTY; without even 
+     the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR 
+     PURPOSE.  See the above copyright notices for more information.
+
+=========================================================================*/
+#ifndef __itkMeanImageFilter_h
+#define __itkMeanImageFilter_h
+
+#include "itkBoxImageFilter.h"
+#include "itkImage.h"
+#include "itkNumericTraits.h"
+
+namespace itk
+{
+/** \class MeanImageFilter
+ * \brief Applies an averaging filter to an image
+ *
+ * Computes an image where a given pixel is the mean value of the
+ * the pixels in a neighborhood about the corresponding input pixel.
+ *
+ * A mean filter is one of the family of linear filters.  
+ *
+ * \sa Image
+ * \sa Neighborhood
+ * \sa NeighborhoodOperator
+ * \sa NeighborhoodIterator
+ * 
+ * \ingroup IntensityImageFilters
+ */
+template <class TInputImage, class TOutputImage>
+class ITK_EXPORT MeanImageFilter :
+    public BoxImageFilter< TInputImage, TOutputImage >
+{
+public:
+  /** Extract dimension from input and output image. */
+  itkStaticConstMacro(InputImageDimension, unsigned int,
+                      TInputImage::ImageDimension);
+  itkStaticConstMacro(OutputImageDimension, unsigned int,
+                      TOutputImage::ImageDimension);
+
+  /** Convenient typedefs for simplifying declarations. */
+  typedef TInputImage InputImageType;
+  typedef TOutputImage OutputImageType;
+
+  /** Standard class typedefs. */
+  typedef MeanImageFilter Self;
+  typedef BoxImageFilter< InputImageType, OutputImageType> Superclass;
+  typedef SmartPointer<Self> Pointer;
+  typedef SmartPointer<const Self>  ConstPointer;
+
+  /** Method for creation through the object factory. */
[…]

### 2007-08-02 itkNoiseImageFilter.h
implement NoiseImageFilter as a subclass of BoxImageFilter
darcs-hash:20070802081914-2fc9d-a8159f5ca68c531b4941f026db73e8a24177b1e3.gz

@@ -1,0 +1,116 @@
+/*=========================================================================
+
+  Program:   Insight Segmentation & Registration Toolkit
+  Module:    $RCSfile: itkNoiseImageFilter.h,v $
+  Language:  C++
+  Date:      $Date: 2006/03/30 15:36:24 $
+  Version:   $Revision: 1.3 $
+
+  Copyright (c) Insight Software Consortium. All rights reserved.
+  See ITKCopyright.txt or http://www.itk.org/HTML/Copyright.htm for details.
+
+     This software is distributed WITHOUT ANY WARRANTY; without even 
+     the implied warranty of MERCHANTABILITY or FITNESS FOR A PARTICULAR 
+     PURPOSE.  See the above copyright notices for more information.
+
+=========================================================================*/
+#ifndef __itkNoiseImageFilter_h
+#define __itkNoiseImageFilter_h
+
+#include "itkBoxImageFilter.h"
+#include "itkImage.h"
+#include "itkNumericTraits.h"
+
+namespace itk
+{
+/** \class NoiseImageFilter
+ * \brief Calculate the local noise in an image.
+ *
+ * Computes an image where a given pixel is the standard deviation of
+ * the pixels in a neighborhood about the corresponding input pixel.
+ * This serves as an estimate of the local noise (or texture) in an
+ * image. Currently, this noise estimate assume a piecewise constant
+ * image.  This filter should be extended to fitting a (hyper) plane
+ * to the neighborhood and calculating the standard deviation of the
+ * residuals to this (hyper) plane.
+ *
+ * \sa Image
+ * \sa Neighborhood
+ * \sa NeighborhoodOperator
+ * \sa NeighborhoodIterator
+ * 
+ * \ingroup IntensityImageFilters
+ */
+template <class TInputImage, class TOutputImage>
+class ITK_EXPORT NoiseImageFilter :
+    public BoxImageFilter< TInputImage, TOutputImage >
+{
+public:
+  /** Extract dimension from input and output image. */
+  itkStaticConstMacro(InputImageDimension, unsigned int,
+                      TInputImage::ImageDimension);
+  itkStaticConstMacro(OutputImageDimension, unsigned int,
+                      TOutputImage::ImageDimension);
+
+  /** Convenient typedefs for simplifying declarations. */
+  typedef TInputImage InputImageType;
+  typedef TOutputImage OutputImageType;
+
+  /** Standard class typedefs. */
[…]

### 2007-08-02 article/Article.tex
some extensions to the article
darcs-hash:20070802141347-2fc9d-b5b8e1950a02fbb3c99701598da470077a334bfc.gz

@@
 
+\subsection{Common filters}
+
+All the filters provided with those packages are implemented as subclasses
+of one of two new classes: {\em itk::BoxImageFilter} and {\em itk::KernelImageFilter}.
+{\em itk::BoxImageFilter} is a subclass of {\em itk::ImageToImageFilter} and provides
+some features used in all the box based filters:
+\begin{itemize}
+  \item the management of the radius of the box,
+  \item and the enlargement of the input requested region, according to the box size.
+\end{itemize}
+{\em itk::KernelImageFilter} is a subclass of {\em itk::BoxImageFilter}, and provides
+some features used in all the kernel based filters, in addition to the features of the
+{\em itk::BoxImageFilter} class:
+\begin{itemize}
+  \item the management of the kernel,
+  \item a redefined SetRadius() method to create a box kernel with the provided radius,
+  \item a default box kernel of radius $1$\footnote{This feature should help many ITK
+beginers who are often {\em not} providing a structuring element to the morphology filters,
+and thus get difficultly understandable crashes.}.
+\end{itemize}
+
+A third filter, {\em itk::SeparableImageFilter}, is used to easily implement a separable
+version of an algorithm, by using its basic implementation in a subclass of
+{\em itk::BoxImageFilter}, or any other subclass of {\em ImageToImageFilter} with
+a {\em SetRadius()} method.
+
+Some filters already in ITK have been modified to use {\em itk::BoxImageFilter} or
+{\em itk::KernelImageFilter}\footnote{itk::MedianImageFilter, itkMeanImageFilter,
+itk::NoiseImageFilter and itk::MorphologyImageFilter.}. Their API is not modified.
+
 \subsection{Morphology filters -- consolidatedMorphology}
@@
 	be attempted. Caution is advised because this is not well tested yet.
+  \item \verb$Cross()$ produces a cross structuring element. It takes the radius 
+	of the structuring element as parameter.
+  \item \verb$Annulus()$ produces an annulus structuring element. It takes the radius 
+	of the structuring element as parameter, as well as the thickness of the annulus,
+	and a boolean to indicate whether the center neighbors should be set inside the
+	neighborhood or not\footnote{The {\em Annulus()} constructor  is a contribution
+of Dan Mueller.}.
   \item \verb$FromImage()$ produces a structuring element from an image. The image is passed
@@
 sliding window infrastructure, which need to be consolidated.
-
-All the kernel based filters in ITK would also benefit to hinerit a common
-super class which manage the neighborhood size and/or the kernel, and which
-provide the infrastructure for the separable operations.
 

### 2007-08-07 article/Article.tex
add threading perf for the fastRankMean filters and update consolidatedMorphology perf2D results to include separable basic algorithm, and use the  right name for the iterative alogrithm
darcs-hash:20070807151858-2fc9d-d60c71adc0bf995f6a546a1404c8b2b2a6e3295d.gz

@@
 \subfigure[Median]{\includegraphics[scale=0.62]{fastRankMean-perfMedian}}
-\caption{Execution times for several algorithms of mean filters (a) and
-median filters (b).\label{fig:perfRank}
-Results are reported in Tables \ref{tab:perfRankMean} and  \ref{tab:perfRankMedian}.}
+\subfigure[Threads]{\includegraphics[scale=0.62]{fastRankMean-perf_threads-large}}
+\caption{Execution times for an increasing kernel size with several algorithms of mean filters (a) and
+median filters (b). Excution times with an an increasing number of threads (c).\label{fig:perfRank}
+Results are reported in Tables \ref{tab:perfRankMean}, \ref{tab:perfRankMedian}
+and \ref{tab:perfRankThreads}.}
 \end{center}
@@
 \centering
-\begin{tabular}{ccccccc}
+\small
+\begin{tabular}{cccccccc}
 \hline
-Radius	&	Basic	&	Separable	&	Histogram	&	Separable histogram	&	Anchor	&	van Herk / Gil Werman	\\
+Radius	&	Basic	&	Iterative	&	Separable	&	Histogram	&	Separable histogram	&	Anchor	&	van Herk / Gil Werman	\\
 \hline
-1	&	0.00477	&	0.00491	&	0.013	&	0.012	&	0.0144	&	0.00725	\\
-2	&	0.0114	&	0.0101	&	0.0161	&	0.0119	&	0.0135	&	0.00743	\\
-3	&	0.024	&	0.0146	&	0.0193	&	0.0116	&	0.0129	&	0.00794	\\
-4	&	0.0444	&	0.0192	&	0.0231	&	0.0114	&	0.0126	&	0.00755	\\
-5	&	0.0714	&	0.0247	&	0.0261	&	0.0114	&	0.0121	&	0.0074	\\
-6	&	0.107	&	0.0299	&	0.0313	&	0.0116	&	0.0127	&	0.00771	\\
-7	&	0.153	&	0.034	&	0.0358	&	0.0113	&	0.0121	&	0.00859	\\
-8	&	0.211	&	0.0407	&	0.038	&	0.0112	&	0.0121	&	0.00773	\\
-9	&	0.281	&	0.0442	&	0.0434	&	0.0114	&	0.012	&	0.00779	\\
-10	&	0.365	&	0.0486	&	0.0446	&	0.0111	&	0.0118	&	0.00763	\\
-15	&	1.03	&	0.074	&	0.0604	&	0.0115	&	0.0116	&	0.00794	\\
-20	&	2.23	&	0.104	&	0.079	&	0.0113	&	0.0113	&	0.00762	\\
-25	&	4.11	&	0.129	&	0.0968	&	0.0114	&	0.0116	&	0.0079	\\
-30	&	6.82	&	0.154	&	0.111	&	0.0114	&	0.0115	&	0.00781	\\
-40	&	15.4	&	0.214	&	0.153	&	0.0116	&	0.0116	&	0.00785	\\
-50	&	29.3	&	0.285	&	0.19	&	0.0119	&	0.0117	&	0.00804	\\
-100	&	226	&	0.767	&	0.355	&	0.0129	&	0.0117	&	0.00809	\\
+1	&	0.00437	&	0.00436	&	0.00734	&	0.00987	&	0.0121	&	0.0111	&	0.00715	\\
+2	&	0.0118	&	0.00872	&	0.00901	&	0.0131	&	0.0121	&	0.0101	&	0.00733	\\
+3	&	0.0241	&	0.013	&	0.0108	&	0.0161	&	0.0118	&	0.00979	&	0.00756	\\
+4	&	0.0446	&	0.0175	&	0.0127	&	0.0213	&	0.0117	&	0.00922	&	0.00736	\\
+5	&	0.0718	&	0.0219	&	0.0149	&	0.0237	&	0.0116	&	0.00923	&	0.00768	\\
+6	&	0.107	&	0.0262	&	0.017	&	0.0262	&	0.0114	&	0.00906	&	0.00768	\\
+7	&	0.154	&	0.0311	&	0.0192	&	0.0315	&	0.0113	&	0.00899	&	0.00779	\\
+8	&	0.211	&	0.0349	&	0.0223	&	0.0334	&	0.0115	&	0.00865	&	0.00759	\\
+9	&	0.281	&	0.0395	&	0.0246	&	0.0385	&	0.0113	&	0.00892	&	0.00784	\\
+10	&	0.365	&	0.0441	&	0.0273	&	0.0399	&	0.0113	&	0.0086	&	0.00767	\\
+15	&	1.03	&	0.0668	&	0.0451	&	0.0585	&	0.0115	&	0.00851	&	0.00798	\\
+20	&	2.23	&	0.0895	&	0.0657	&	0.0761	&	0.0118	&	0.00819	&	0.0077	\\
+25	&	4.11	&	0.114	&	0.0911	&	0.0958	&	0.0114	&	0.00826	&	0.00778	\\
+30	&	6.84	&	0.139	&	0.121	&	0.108	&	0.0115	&	0.00847	&	0.00801	\\
+40	&	15.5	&	0.194	&	0.195	&	0.145	&	0.0116	&	0.00872	&	0.00811	\\
+50	&	29.4	&	0.257	&	0.283	&	0.185	&	0.0118	&	0.00858	&	0.00798	\\
+100	&	227	&	0.719	&	0.974	&	0.348	&	0.0127	&	0.00858	&	0.0082	\\
 \hline
@@
 \centering
+\small
 \begin{tabular}{ccccc}
@@
 \centering
+\small
 \begin{tabular}{cccc}
 \hline
-Radius	&	Histogram	&	van Herk / Gil Werma	&	Anchorn	\\
+Radius	&	Histogram	&	van Herk / Gil Werman	&	Anchorn	\\
 \hline
@@
 \centering
+\small
+\begin{tabular}{ccccccc}
+\hline
+Threads	&	fastApproxMaskRank	&	fastApproxRank	&	maskedRank	&	movingWindowMean	&	rank	&	separableMean	\\
+\hline
+1	&	1.83	&	1.78	&	3.99	&	2.17	&	4.36	&	1.23	\\
+2	&	0.921	&	0.898	&	2.02	&	1.09	&	2.21	&	0.634	\\
+3	&	0.698	&	0.704	&	1.48	&	0.788	&	1.48	&	0.471	\\
+4	&	0.516	&	0.505	&	1.07	&	0.559	&	1.12	&	0.403	\\
+5	&	0.613	&	0.637	&	1.24	&	0.764	&	1.34	&	0.461	\\
+6	&	0.582	&	0.548	&	1.12	&	0.607	&	1.17	&	0.42	\\
+7	&	0.546	&	0.539	&	1.14	&	0.642	&	1.26	&	0.398	\\
+8	&	0.513	&	0.506	&	1.06	&	0.598	&	1.13	&	0.402	\\
+9	&	0.569	&	0.558	&	1.12	&	0.615	&	1.22	&	0.412	\\
+10	&	0.558	&	0.55	&	1.07	&	0.603	&	1.16	&	0.408	\\
+\hline
+\hline
+\end{tabular}
+\caption{Execution times in seconds for median and mean filters using direct and sliding window implementations applied to the $3200 \times 2400$ image.\label{tab:perfRankThreads}}
+\end{table}
+
+\begin{table}[phtb]
+\centering
+\small
 \begin{tabular}{ccccc}
@@
 \hline
+\small
 \end{tabular}

### 2007-10-26 seminar/seminar.tex
Added seminar sources to repository
darcs-hash:20071026003600-fafb9-0e4794d7ff66b5e4aa95e6a729934d0e8fead104.gz

\documentclass[pdf,prettybox]{prosper}
\usepackage{graphicx}
\title{Efficient implementation of kernel filtering}
\author{Richard Beare{$^1$} {\small{and}} Ga\"etan Lehmann{$^2$}}
\institution{{$^1$}Department of Medicine, Monash University, Australia.\\
{$^2$}INRA, UMR 1198; ENVA; CNRS, FRE 2857, Biologie du D\'eveloppement et
Reproduction, Jouy en Josas, F-78350, France}

\Logo(1cm,-1cm){\includegraphics[height=0.5cm]{monash}\hspace{2cm}\includegraphics[height=0.5cm]{LogoINRA-Couleur}\hspace{2cm}\includegraphics[height=0.5cm]{mima2}}

\begin{document}
\maketitle

\begin{slide}{Introduction}
\begin{itemize}
\item Describe methods that reduce computational complexity of kernel filters by orders of magnitude.
\item Discuss the tradeoffs that need to be considered when using these methods.
\end{itemize}
\end{slide}

\begin{slide}{Motivation}
Large median filter $111 \times 111$ for background estimation in microscopy image.
\begin{itemize}
\item Direct implementation takes at least 20 minutes.
\item Efficient approximation takes less than 5 seconds.
\end{itemize}
\end{slide}

\overlays{5}{%
\begin{slide}{Improving filter performance}
\begin{itemstep}
\item Recursive computation
	\begin{itemize}
	\item An output value can be computed based on a neighbouring output.
	\item This reduces complexity by eliminating repeated computation.
	\end{itemize}
\item Separability
	\begin{itemize}
	\item Certain multidimensional kernels can be decomposed into a number of one dimensional operations.
	\end{itemize}
\end{itemstep}
\end{slide}
}

\begin{slide}{Improving filter performance}
\begin{itemize}
\item ITK Gaussian filtering classes use both approaches.
\end{itemize}
\end{slide}

\overlays{2}{%
\begin{slide}{Sliding window}
\onlySlide*{1}{%
\begin{centering}
\begin{figure*}[p]
\includegraphics[width=6cm]{sliding}
\end{figure*}
\end{centering}}%
\onlySlide*{2}{
\begin{centering}
\begin{figure*}[p]
\includegraphics[width=6cm]{slidingB}
\end{figure*}
\end{centering}}%
\end{slide}
}

\begin{slide}{Sliding window}
\begin{itemize}
\item Collect a summary of all pixels in the neighbourhood.
\item Compute the output value from the summary.
\item Move the window and update the summary.
\end{itemize}
\end{slide}

\begin{slide}{Sliding window}
Reduces complexity from $O(n^d)$ to $O(n^{d-1})$

The standard updatable summaries are:
\begin{itemize}
\item Sum or sum of squares for mean and variance computations.
\item Histograms for rank filters including mathematical morphology.
\end{itemize}
\end{slide}

\overlays{3}{%
\begin{slide}{Separability/Decomposition}
\begin{itemstep}
\item Some useful filters can be computed using a cascade of operations along lines.
\item Special forms of updateable computations exist for some operations along lines.
\item These lead to operations with complexity independent of line length.
\end{itemstep}
\end{slide}
}
\begin{slide}{Mathematical Morphology}
\begin{itemize}
\item van Herk/Gil Werman algorithm.
\item Erosions/dilations along lines with 3 operations per pixel.
\end{itemize}
\end{slide}

\begin{slide}{Mathematical Morphology}
\vspace{0.4cm}
\begin{centering}
\begin{figure*}
\includegraphics[width=7.5cm]{vHGWexpl}
\end{figure*}
\end{centering}
\end{slide}

\begin{slide}{Mathematical Morphology}
\begin{itemize}
\item Decomposition of shapes as follows:
	\begin{itemize}
	\item Hyper-boxes - cascade of line operations parallel to image axes.
	\item Appoximate circles in 2D.
	\item Experimental approximation to spheres.
	\end{itemize}
\end{itemize}
\end{slide}

\begin{slide}{Mathematical Morphology}
\vspace{0.4cm}
\begin{centering}
\begin{figure*}
\includegraphics[width=4cm]{kernel}
\end{figure*}
\end{centering}
\end{slide}

\begin{slide}{Robust smoothing filters}
\begin{itemize}
\item Median filters aren't separable.
\item Applying a cascade of line based rank filters produces a fast, robust smoother.
\end{itemize}
\end{slide}

\begin{slide}{Accumulation image methods}
\begin{itemize}
\item Useful for sums, means or variances of boxes.
\item One accumulation image can be used to evaluate lots of different kernel sizes.
\item Can also be used for binary morphology and binary median filters.
\end{itemize}
\end{slide}

\begin{slide}{Accumulation image methods}
\vspace{0.4cm}
\begin{centering}
\begin{figure*}
\includegraphics[width=7cm]{accum2D}
\end{figure*}
\end{centering}
\end{slide}

\begin{slide}{Accumulation image methods}
\vspace{0.4cm}
\begin{centering}
\begin{figure*}
\includegraphics[width=7cm]{accumMean2d}
\end{figure*}
\end{centering}
\end{slide}

\begin{slide}{Summary of filters}
\begin{itemize}
\item {\bf Mathematical Morphology using flat structuring elements} - arbitrary, box, polygon.
\item {\bf Accumulation image methods - mean and standard deviation}. Only one kernel size.
\item {\bf Rank filters} - arbitrary, boxes - separable approximation, also masked versions.
\end{itemize}
\end{slide}

\begin{slide}{Summary of complexities}
\begin{itemize}
\item Sliding window - $O(n^{d-1})$ - arbitary morphology, arbitary rank and mean filters.
\item Box convolution - constant complexity.
\item MM along lines - constant complexity - approximate circles, boxes.
\item Rank operations along lines - uses sliding windows and provides a fast, robust smoother.
\end{itemize}
\end{slide}

\end{document}
